---
title: "SOC 301-01: Social Statistics"
author: "Chester Ismay"
output:
  ioslides_presentation:
    incremental: no
    logo: ../figs/boxer.jpg
    fig.align: center
    widescreen: true
subtitle: Five Main Verbs + `%>%` (Practice)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center')
```


## Learning Quote of the Day {.flexbox .vcenter}

> "It’s not just what you know, but how you practice what you know that determines how well the learning serves you later."

― Peter C. Brown, **Make It Stick**

## Important Concept: Piping

Piping allows you to take the output of one function and **pipe** it as the input of
the next function. You can string along several pipes to form a single **chain**.

* R Command: `%>%`
* Described as: "_and then_". 



## Important Concept: Piping

Ex:  say you want to apply functions `h()` and `g()` and then `f()` on data `x`.  You can do

* `f(g(h(x)))` OR
* `x %>% h() %>% g() %>% f()`

This

> * saves you from confusing nested parentheses
> * emphasizes the **sequential** breaking down of tasks, making it more readable
> * i.e. Do this _then_ do this _then_ do this _then_ ...




## Important Concept: Piping

Pipes are always directed to the first argument of any function. The following three bits of R code
do the same thing, extract all january flights:

```{r,eval=FALSE, echo=TRUE}
library(nycflights13)
data(flights)

# Bit 1: No piping
filter(flights, month == 1)

# Bit 2: Piping. Note no comma
flights %>% filter(month == 1)

# Bit 3: Piping across multiple lines (preferred for legibility)
flights %>% 
  filter(month==1)
```