---
title: "SOC 301-01: Social Statistics"
author: "Chester Ismay"
output:
  ioslides_presentation:
    incremental: no
    logo: ../figs/boxer.jpg
    fig.align: center
    widescreen: true
subtitle: Five Main Verbs + `%>%` (Practice)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center')
lc <- 0
library(ggplot2)
```

<style type="text/css">
    ol { list-style-type: upper-alpha; }
</style>

## Learning Quote of the Day {.flexbox .vcenter}

> "It’s not just what you know, but how you practice what you know that determines how well the learning serves you later."

― Peter C. Brown, **Make It Stick**

# Plicker time!

## Problem `r lc <- lc + 1; lc`

What is faceting?

a. Creates small multiples of the same plot over a different categorical variable
b. Creates small multiples of the same plot over a different numerical variable
c. Create large multiples of the same plot over a different categorical variable
d. Create small variables of the same plot over a similar numerical variable

## Problem `r lc <- lc + 1; lc`

What type of plot is usually preferred for an explanatory categorical variable and a response categorical variable?

a. Faceted boxplot 
b. Side-by-side barplot
c. Stacked barplot
d. Faceted barplot

## Problem `r lc <- lc + 1; lc`

Given the 'weather' dataset, what function would we use to choose only the variables of humidity and precipitation?

a. `arrange(weather, humid, precip)`
b. `select(weather, humid, precip)`
c. `filter(weather, humid, precip)`
d. `summarize(weather, humid)`

## Problem `r lc <- lc + 1; lc`

What is the process of decomposing frames into less redundant tables without losing info?

a. framing
b. decomposing
c. tabling
d. normalizing

## Problem `r lc <- lc + 1; lc`

What R code would you use to find the mean and standard deviation of temperature in the weather data set?

a. 

    ```{r eval=FALSE}
    weather %>% summarize(mean = mean(temp),
              std_dev = sd(temp))
    ```

b. 

    ```{r eval=FALSE}
    weather %>% summarize(mean = mean(temp, na.rm = TRUE),
              std_dev = sd(temp, na.rm = TRUE))
    ```

c. 

    ```{r eval=FALSE}
    weather %>% summarize(mean = (mean = temp),
              std_dev = (sd = temp))
    ```

d. 

    ```{r eval=FALSE}
    weather %>% summarize(mean = (mean == temp),
              std_dev = (sd == temp))
    ```

## Problem `r lc <- lc + 1; lc`

Find the mean `gdpPercap` of each region for each year 

a. 

    ```{r eval=FALSE}
region_perCap <- gap %>% group_by(year) %>%
  summarize(mean_perCap = mean(gdpPercap))
```

b. 

    ```{r eval=FALSE}
    region_perCap <- gap %>% 
      summarize(region, year, mean_perCap = mean(gdpPercap))
    ```

c.

    ```{r eval=FALSE}
    region_perCap <- gap %>% group_by(region, year) %>% 
      summarize(mean_perCap = mean(gdpPercap))
    ```

d. 
    ```{r eval=FALSE}
    region_perCap <- gap %>% group_by(region) %>% 
      summarize(mean_perCap = mean(gdpPercap))
    ```

## Reflect

- What concepts are you having the hardest time with right now?  Be as specific as possible.

- What else can I do to help you work with `dplyr` and `ggplot2`?

- What else will you be doing to help yourself work with `dplyr` and `ggplot2`?

