---
title: "All practice problems from PS9"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r include=FALSE}
lc <- 0
```

***

# General Questions

# Problem `r lc <- lc + 1; lc`

Which of the following is NOT one of the five main verbs?

A. Select
B. Filter
C. Dplyr
D. Mutate

# Problem `r lc <- lc + 1; lc`

Which of the following is correct about the five main verbs?

A. Select- Pick rows based on conditions about their values
B. Arrange- Sort the rows based on one or more variables
C. Mutate- Create summary measures of variables (or groups of observations on variables using group_by)
D. None of the above

# Problem `r lc <- lc + 1; lc`

What is this %>% called?

A. Rename
B. It does not have a name
C. Filter
D. Pipe

# Problem `r lc <- lc + 1; lc`

Which of the following is true?

A. < corresponds to “greater than”
B. > corresponds to “less than”
C. >= corresponds to “greater than or equal to”
D. All of the above 

# Problem `r lc <- lc + 1; lc`

What is the correct way to rename?

A. "New before, old after"
B. "Old before, new after"
C. "Just new before"
D. It doesn't matter

# Problem `r lc <- lc + 1; lc`

What are the five main verbs used in manipulating data?

# Problem `r lc <- lc + 1; lc`

How can I select a range of columns from a data set?

# Problem `r lc <- lc + 1; lc`

How do I select columns starting with specific letters?

# Problem `r lc <- lc + 1; lc`
What do these symbols correspond to: >, <, >=, <=, and !=

# Problem `r lc <- lc + 1; lc`

What does the 'count' function do?

# Problem `r lc <- lc + 1; lc`

Why is the `select` function an important tool when analyzing data frames?

A. Allows you to select specific variables in a data frame.

B. Alters existing data frame or creates new data frame within existing data.

C. An easier way to extract variables that you want to analyze in a data frame

D. All the above

# Problem `r lc <- lc + 1; lc`

The ____ operator grabs columns and the ____ operator grabs rows. 

a. Select, Filter 
b. Select, Observe
c. Variable, Filter 
d. Variable, Observe


# Problem `r lc <- lc + 1; lc`

What is this `%>%` and what does it do? 

a. This is called the pipe and it is to chain dplyr functions. 

b. This is called the pipe and it is to convert all numerical data in the function to percentages. 

c. This has no name and it is to chain dplyr functions. 

d. This has no name and it is to covert all numerical data in the function to percentages. 

# Problem `r lc <- lc + 1; lc`

What does the `select` feature do?

# Problem `r lc <- lc + 1; lc`

What does the function `inner_join` do?

A. `inner join` brings two different data frames together. 
B. `inner join` takes two data frames that have been joined together and separates them.
C. `inner join` takes a group of numbers and averages them in order to bring them together. 
D. `inner join` does nothing in R. 

# Problem `r lc <- lc + 1; lc`

 What are the five MAIN verbs when using `dplyr`?

A. `select`, `filter`, `group_by`, `mutate`, `summarize`
B. `filter`, `plot`, `mutation`, `choose`, `pick`
C. `select`, `filter`, `summarize`, `mutate`, `arrange`
D. `library`, `data`, `install`, `mutate`, `organize`

# Problem `r lc <- lc + 1; lc`

Which of the following shows the correct definition of the given operations? (choose all the apply)

A. `!` corresponds to "not equal to"
B. `|` corresponds to "or"
C. `>` corresponds to "greater than"
D. `%>%` chains together functions in the `dyplr` package 
E. All of the above show the correct definition for the provided operation 

# Problem `r lc <- lc + 1; lc`

What does the `?` operator do in R?

A. `?` followed immediately by a function (i.e. `?summarize`) will provide you with information about the summarize function in the `dyplr` package when run in your environment section. 
B. `?` followed immediately by a `+` and the a function (i.e. `?summarize`) will provide you with information about the summarize function in the `dyplr` package.
C. a function followed immediately by `?` (i.e. `summarize?`) will provide you with information about the summarize function in the `dyplr` package when run in your console. 
D. `?` followed immediately by a function (i.e. `?summarize`) will provide you with information about the summarize function in the `dyplr` package when run in your console. 

***

# Non-`gap` practice problems

# Problem `r lc <- lc + 1; lc`

Using the `flights` data set, create a table that isolates only variables that end with "time".

# Problem `r lc <- lc + 1; lc`

Using the `flights` data set, create a table that focuses just on flights from New York City to Los Angeles. 

# Problem `r lc <- lc + 1; lc`

Create a small table that summarizes the mean and standard deviation of `wind_speed` for each month. 

# Problem `r lc <- lc + 1; lc`

Use the `arrange` function to show the most frequent carriers. 


# Problem `r lc <- lc + 1; lc`

Which of the following codes is correct? (Using the dyplr package)

```{r A., eval=FALSE}
ten_freq_dests -> flights %>%
  group_by(dest) %>%
  summarize(num_flights = n() %>%
  top_n = 10) %>%
  arrange(desc(num_flights)
```

```{r B., eval=FALSE}
ten_freq_dests <- flights %>%
  group_by(dest) %>%
  summarize(num_flights = n()) %>%
  top_n(n = 10) %>%
  arrange(desc(num_flights))
```

```{r C., eval=FALSE}
ten_freq_dests <- flights +
  group_by(dest) %>%
  summarize(num_flights = n()) +
  top_n(n = 10) +
  arrange(desc(num_flights))
```

```{r D., eval=FALSE}
ten_freq_dests <- flights, 
  group_by(dest), 
  summarize(num_flights = n()),
  top_n(n = 10),
  arrange(desc(num_flights))
```

# Problem `r lc <- lc + 1; lc`

How would you combine all the different levels of college attendance in "profiles" into the more managable categories 'college', 'some college', 'no college'?

# Problem `r lc <- lc + 1; lc`

Select 'body type', 'age', and 'diet' from the "profiles" dataframe.

# Problem `r lc <- lc + 1; lc`

Filter the "profiles" dataframe by profiles of men betweent the ages of 21-24.

# Problem `r lc <- lc + 1; lc`

Calculate the mean age of people with OkCupid acocunts.

# Problem `r lc <- lc + 1; lc`

Select the profiles of everyone who is not 'straight'.

***

# `gap` Practice Problems

```{r load_gap}
gap <- read.csv("/shared/isma5720@pacificu.edu/gapminder.csv", 
                      stringsAsFactors = FALSE)
```


# Problem `r lc <- lc + 1; lc`

What is the standard deviation for population in 1972?

# Problem `r lc <- lc + 1; lc`

How would you pipe together population and life expectancy in 2007?

# Problem `r lc <- lc + 1; lc`

What is mean for gdpPercap 1972?

# Problem `r lc <- lc + 1; lc`

Use the filter function for lifeExp?

# Problem `r lc <- lc + 1; lc`

What is the mean for dem_rank in 1952?

# Problem `r lc <- lc + 1; lc`

Which subRegion(s) in 1992 were ‘middle of the road’ in their dem_rank?

# Problem `r lc <- lc + 1; lc`

How many countries end with the letter ‘a’, and what are their names?

# Problem `r lc <- lc + 1; lc`

Which subRegion(s) have a life expectancy above 25 but below 50?

# Problem `r lc <- lc + 1; lc`

Which countries in 1992 were strongly autocratic?

# Problem `r lc <- lc + 1; lc`

What is the mean population for Cambodia from the year 1952 to 1982?

# Problem `r lc <- lc + 1; lc`

List all strongly democratic countries from the year 1992.

# Problem `r lc <- lc + 1; lc`

Name the countries with the top 10 highest life expectancies from 1957.

# Problem `r lc <- lc + 1; lc`

What are the top 5 countries with the biggest population increases from 1957-2007?

# Problem `r lc <- lc + 1; lc`

Name the country with the highest life expectancy from each subRegion.

# Problem `r lc <- lc + 1; lc`

What is the mean gdpPercap for each region in 1957?

# Problem `r lc <- lc + 1; lc`

Which subregions had the 5 highest populations in 2007?

# Problem `r lc <- lc + 1; lc`

Plot 'lifeExp' by 'region' and fill the plot by 'subregion'.

# Problem `r lc <- lc + 1; lc`

What is the mean 'gdpPercap' in the 'gap' data set?

# Problem `r lc <- lc + 1; lc`

Plot 'lifeExp' over 'dem_rank' for the year 2007.

# Problem `r lc <- lc + 1; lc`

Extract only variables 'year', 'lifeExp', and 'pop'.

# Problem `r lc <- lc + 1; lc`

Create a data frame consisting only of data from 1952. 

# Problem `r lc <- lc + 1; lc`

Find the mean life expectancy by each region in 1952 with the highest mean expectency at the top of the list. 

# Problem `r lc <- lc + 1; lc`

Find the combined mean GDP per capita for 1952 and 2007. 

# Problem `r lc <- lc + 1; lc`

Create data frame showing the data from the years 1952 and 2007 arranged in descensing order by gdpPercap. 
Extra- Now select just the highest 25 percapita gdps, still in descending order. 

# Problem `r lc <- lc + 1; lc`

Create a new data frame showing total GDP of each country in each year. 
Extra - reorder the columns so that GDP, gdpPerCap and pop appear in that order at the end of the data set but before `dem_rank`


# Problem `r lc <- lc + 1; lc`

Filter out the years 1987 and before then. Only show the countries that correspond to 1992 or after that. 

a. `nintys_to_thousands <- filter(gap, !(year<=1987))`
b. `nintys_to_thousands <- select(gap, !(year<==1987))`
c. `nintys_to_thousands <- filter(gap, (year<=1987))`
d. `nintys_to_thousands <- select(gap, !(year<=1987))`

# Problem `r lc <- lc + 1; lc`

Collect the mean and the standard deviation of all the populations. 

a. `unt1 <- summarize(gap, mean= mean(pop, na.rm=TRUE), std_dev = sd(pop, na.rm=TRUE))`

b. `unt1 <- summarize(gap, mean= mean"pop", na.rm=TRUE), std_dev = sd("pop", na.rm=TRUE)`

c. `unt1 <- summarize(gap, mean= mean(pop), std_dev = sd(pop))`

d. `unt1 <- summarize(gap, mean= mean"pop"), std_dev = sd("pop")`


# Problem `r lc <- lc + 1; lc`

What does the following R code produce?

```{r P1, eval=FALSE}
gap_new <- select(gap, country, lifeExp)
gap
```

A. A data set with `country` and `lifeExp`.

B. A data frame that outputs all the variables in the gap data frame except for `country` and `lifExp`. 

C. A new data set with only `country` and `lifeExp` as the variables observed. 

D. None of the above



# Problem `r lc <- lc + 1; lc`

If you want to get the average of the `lifExp` variable and it's standard deviation while omitting NA variables, which R chunk would you use?

A.
```{r P3A, eval=FALSE}
gap <- summary(gap,
          mean = mean(lifExp),
          std_dev = sd(lifExp)
          )
```

B.
```{r P3B, eval=FALSE}
gap_summary <- summarize(gaplifExp,
          mean = mean(lifExp, na.rm = TRUE),
          sd = (lifExp)
          )
gap
```

C.
```{r P3C, eval=FALSE}
gap <- summarize(gap,
          mean = avg(lifeExp, na.rm = FALSE),
          std_dev = sd(pop, na.rm = TRUE)
          )
gap
```

D. 
```{r P3D, eval=FALSE}
gap <- summarize(gap,
          mean = mean(lifeExp, na.rm = TRUE),
          std_dev = sd(lifeExp, na.rm = TRUE)
          )
gap
```

# Problem `r lc <- lc + 1; lc`

What does the `std_dev`, or standard deviation of a data frame represent?

A. Deviation from the mode of the data frame.

B. A spread of the values observed in a data frame that accounts for how far the range of values are from the mean value.

C. On average how far apart the values in the data set are.

D. None of the above.

# Problem `r lc <- lc + 1; lc`

What is the importance of the pipe function ` %>% ` when examining the data frame in the chunk below:

```{r P5, eval=FALSE}
gap %>%
  top_n(n = 10, wt = lifeExp)
```

A. Select data in `lifeExp`

B. Shorter and less confusing code to extract the data from the first ten observations for the lifeExp variable.

C. Create new column in data frame titled `top_n)

D. All the above

# Problem `r lc <- lc + 1; lc`

How would you is isolate just Afganistan?

# Problem `r lc <- lc + 1; lc`

What would gap_info <- select(gap, year:pop, gdpPercap:dem_rank) gap_info do?

# Problem `r lc <- lc + 1; lc`

What is the standard deviation of gdpPercap?

# Problem `r lc <- lc + 1; lc`

What would you need to put to arrange the life expectancy into a smallest to largest?

# Problem `r lc <- lc + 1; lc`

How would you get the five number summary of gdpPercap?



