---
title: "Lab 1 - Intro to R and RStudio"
author: "Key"
date: "week 1"
output: 
  html_document: 
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(oilabs)
```

* * *

## Lab report

#### Exercise 1: 

```{r eval = FALSE}
arbuthnot$girls
```

#### Exercise 2:

As you can see in the plot below, the number of girls steadily rose over time,
though there was a period from 1640 to 1660 when there was a notable drop
in baptisms.

```{r}
qplot(x = year, y = girls, data = arbuthnot, geom = "line")
```

#### Exercise 3:

```{r echo = FALSE}
arbuthnot <- mutate(arbuthnot, total = boys + girls)
arbuthnot <- mutate(arbuthnot, boy_ratio = boys / total)
qplot(x = year, y = boy_ratio, data = arbuthnot, geom = "line")
```

The plot above shows that the proportion of boys varied year to year but stayed
consistently above .5. It should be noted that though the counts appear highly
variable year to year, the y-axis shows that the changes are actually quite
small.

* * *

### On your own:

#### 1:

```{r}
dim(present)
names(present)
```

By viewing the present data set, we can see that it has data from 1940 - 2010.
The dimension are 63 rows and 3 columns. The columns are the same as in the 
`arbuthnot` data set: `year`, `boys`, and `girls`.

#### 2:

These counts are several orders of magnitude larger than those in `arbuthnot`.

#### 3:

```{r}
present <- mutate(present, total = boys + girls)
present <- mutate(present, boy_ratio = boys / total)
qplot(x = year, y = boy_ratio, data = present, geom = "line")
```

The plot above shows that there as been a gradual decline in proportion of births
that are boys, with the exception of an uptick in the 1960s. Again, the scale
on the y-axes is quite zoomed in, so the variation is exaggerated. Similar to
`arbuthnot`, the proportion is consistently above .5.

#### 4:

```{r}
arrange(present, desc(total))
```

The sorted list shows that the 1961 is the year with the greatest number of total
births in the US.

