---
title: "Week Two, Day 1"
author: "Chester Ismay"
date: "January 29, 2016"
output: html_document
---

***

```{r setup, include=FALSE}
# List of useful packages
pkg <- c("dplyr", "ggplot2", "knitr")

# Check if packages are not installed and assign the
# names of the uninstalled packages to the variable new.pkg
new.pkg <- pkg[!(pkg %in% installed.packages())]

# If there are any packages in the list that aren't installed,
# install them
if (length(new.pkg)) {
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
}

# Load the packages into the current environment
library(knitr)
library(dplyr)
library(ggplot2)

# Set number of digits to display
options(digits = 3)
```

1. Plug in audio

1. Tell them I will be gone February to Anchorage (February 13 - February 17)

  - Lectures on Monday, February 15th and Wednesday, February 17th will be in Psych 105
  - Joint meeting with other 141 section
  - Labs will meet as usual on Tuesday with Andrew at the helm

2. Quiz 1 will be on Friday over Chapter 1 of the text
  
  - Start working on PS2 before Friday to help you prepare for the Quiz on Friday

3. Demo Ella's lab

3. **Recap**

  - Correlation does NOT imply <u>causation</u>
  - Causation can only be inferred from <u>randomized</u> experiments (ones that randomly assign treatments to subjects)
  - <u>Random sampling</u> allows us to confidently make inferences about populations through collection of representative samples.
  
4. Today we will be talking about numerical and graphical ways to describe data
  - **Draw plots on the board**

```{r, echo=FALSE, results="asis"}
set.seed(20160201)
gpa <- round(rnorm(10, mean = 3, sd = 0.5), 2)
kable(data.frame(t(gpa)), align = c("c", "c"))
```
  
  - We denote variables with capital letters
    - e.g., $X$ = Cumulative GPA of Reed students
    - What's $\mu$? Based on 2014-15 data, $\mu = 3.15$
  - We denote specific observations of that variable (the sample data) by small letters
    - e.g., $x_1 = `r gpa[1]`$, $x_2 = `r gpa[2]`$, $\cdots$, $x_{10} = `r gpa[10]`$, --- $n = 10$
    
```{r dotplot, echo=FALSE, fig.height=1}
data.frame(gpa) %>% ggplot(aes(x = gpa)) + 
  geom_dotplot(binwidth = 0.05) +
  scale_y_continuous(name = "", breaks = NULL)
```

```{r hist, echo=FALSE, fig.height=2}
data.frame(gpa) %>% ggplot(aes(x = gpa)) + 
  geom_histogram(binwidth = 0.5, colour = "black", fill = "goldenrod")
```

```{r densityplot, echo=FALSE, fig.height=2}
data.frame(gpa) %>% ggplot(aes(x = gpa)) + 
  geom_density()
```

- Area under the curve between any two $x$ values gives the proportion of $y$ values that fall in that given range (integration from calculus)

```{r boxplot, echo=FALSE, fig.height=1}
data.frame(gpa) %>% ggplot(aes(x = '', y = gpa)) + 
  geom_boxplot() +
  coord_flip()
```

5. Go to slides

  - How would we characterize the plots for GPA based on shape?
  - Go over Shape slide

6. Go to Plicker question 1

7. Come back to slides 

  - The Typical Human video
  - go through until Plicker 2

8. Go back to slides

  - Which line corresponds to mean? Which to median?

**Finish reading over Chapter 1, look over your Lab 1 feedback.**  

**We will be discussing describing/manipulating data using the PIPE operator and how to work with data frames in R tomorrow during lab**