---
title: "SOC 301-01: Social Statistics"
author: "Chester Ismay"
output:
  ioslides_presentation:
    incremental: no
    logo: ../figs/boxer.jpg
    fig.align: center
    widescreen: true
subtitle: Data Viz - Histograms and Boxplots
---


## Learning Quote of the Day {.flexbox .vcenter}

> A short study break—-five, ten, twenty minutes to check in on Facebook, respond to a few emails, check sports scores—-is the most effective technique learning scientists know of to help you solve a problem when you’re stuck.

- Benedict Carey, _How We Learn: The Surprising Truth About When, Where, and Why It Happens_

# Plicker time

# Go over the LC questions

## LC4.1 - LC4.4

- Review your responses in partners

## LC4.5 - LC4.9

- Review your responses in partners
    - Someone different than last time

## LC4.10 - LC4.13

- Review your responses in partners
    - Someone different than both of the last two times

# R problems for practice

## Practice problem 1

- Produce a histogram of the humidity recordings in the `weather` data frame with 15 bins, missing values removed, fill with a color of your choosing, and a border color of your choosing.

- Describe in a few sentences how this plot is different from the last plot.

- What more information does this tell you that the previous plot doesn't?

## Practice problem 2 (CHALLENGE)

Produce the following boxplot:

```{r echo=FALSE, fig.height=3.5, fig.width=9.5, fig.pos="center"}
library(nycflights13)
library(ggplot2)
ggplot(weather, aes(x = origin, y = pressure, fill = origin)) + 
  geom_boxplot(na.rm = TRUE, color = "orange") +
  labs(x = "Airport", y = "Atmospheric Pressure", title = "Atmospheric Pressure By Airport in 2013")
```

- What does this plot tell us about how the response variable differs by the different levels of the explanatory variable?

<!--
## Practice problem 5 (CHALLENGE)

Produce the following faceted boxplot:

```{r echo=FALSE, fig.height=3}
library(nycflights13)
library(ggplot2)
ggplot(weather, aes(x = origin, y = temp)) + 
  geom_boxplot(na.rm = TRUE) +
  labs(x = "Airport", y = "Temperature", title = "Temperature By Airport in 2013") +
  facet_wrap(~ month)
```

-->

# Finish Lab 2


## To do for next time

- Finish Lab 2 by 3 PM tomorrow
    - Make sure to let me know who you are working with and where the project is stored
- Read Section 4.4 of [MODERN DIVE textbook](https://ismayc.github.io/moderndiver-book/4-viz.html#barplots)
- Review comments on Lab 2
    - Lab 2 revisions due on Thursday at 3 PM
    - Make sure to Knit to HTML for full credit
    - You must attempt all exercises in the lab for credit.  "I don't know how to do this." or "I'm not sure." are not valid reasons.  I'm here to help, but you have to take the initiative to reach out to me!
- Complete [PS5](https://goo.gl/forms/ZEn8SSVtExPEdK7d2) by 10 AM on Wednesday

## Plans for next time

- Plicker questions over Sections 4.1 - 4.4
- More review of plotting in R with practice problems
- Time to ask questions about Lab 2


## Closing connections

On a half-sheet of paper, 

- Write your name on the front and answer
    - What is the difference between a histogram and a boxplot?

- On the back of this sheet, answer the following questions:

    1. What is the definition of "observational unit"?
    2. What are the three properties of a **tidy** data set?
    3. What are the Five Named Graphs we will explore in this chapter?