---
title: 'NBA Over/Under 2025-2026 Season'
author: Created by Chester Ismay
output: html_document
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path('../2026-nba-over-under.html')) })
---

```{r include=FALSE}
#source("01-03.R")
#Sys.setenv(RSTUDIO_PANDOC = "/Applications/RStudio.app/Contents/MacOS/pandoc")
#Sys.setenv(RSTUDIO_PANDOC = "/usr/local/bin/pandoc")

#remotes::install_github("https://github.com/abresler/nbastatR/commit/f528c9d555ce3baf58e45c6b44318d92faa26b25", force = TRUE)

Sys.setenv("VROOM_CONNECTION_SIZE" = 2e6)

library(httr)
library(jsonlite)
library(tidyverse)
library(magrittr)
library(curl)
library(memoise)
library(furrr)
library(future)
library(glue)
library(stringi)
library(rvest)
library(DT)

library(htmltools)

nbastatR_files <- list.files(path = "nbastatR-r-files",
                             full.names = TRUE)
purrr::walk(nbastatR_files, source)
this_year <- "2025-26"
```

```{r include=FALSE}
r_files1 <- c("01-picks-formatting.R",
              "02a-get-game-scores.R")
purrr::walk(r_files1, source)

reticulate::source_python('02b-get_game_scores.py')
#reticulate::source_python('~/Desktop/ismayc.github.io/nba-over-under-2022-2023/02b-get_game_scores.py')

r_files2 <- c("02c-get-game-scores.R",
              "03-data-wrangling.R")
purrr::walk(r_files2, source)

knitr::opts_chunk$set(echo = FALSE)
library(lubridate)
library(DT)
num_players <- 8
```

Last updated at `r format(Sys.time() - lubridate::hours(8), "%Y-%m-%d %H:%M")` US Pacific time

# {.tabset}

## By team {.tabset}

### By week

<!-- (Skipped until Week 2)  -->

```{r include=FALSE}
first_monday_after_first_week_start <- function(start_date) {
  d <- as.Date(start_date)
  w <- as.POSIXlt(d)$wday          # 0=Sun ... 1=Mon ... 6=Sat
  add <- (1 - w) %% 7              # days until next Monday (could be 0 if Monday)
  if (add == 0) add <- 7L          # ensure *after* the start date
  d + add
}

first_monday <- first_monday_after_first_week_start(nba_season_start_date)
first_monday
```

This plot is updated early each week with the previous week's results (starting on Monday, `r format(first_monday, "%B %d, %Y")
`).

```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE, eval=(Sys.Date() >= first_monday_after_first_week_start(nba_season_start_date))}
weekly_standings <- standings_grid %>% 
  mutate(day_of_week = wday(date, label = TRUE)) %>% 
  filter(day_of_week == "Sun") %>% 
  select(-day_of_week) %>% 
  rename(Date = date, `Team Name` = team)

weekly_team_track <- ggplot(
  weekly_standings,
  aes(x = Date, y = `Current Win %`, label = `Current Record`)
) +
  geom_point() +
  geom_line() +
  geom_segment(aes(x = min(Date), xend = max(Date), 
                   y = `Vegas Win Projection %`, 
                   yend = `Vegas Win Projection %`),
               color = "blue") +
  facet_wrap(~ `Team Name`, nrow = 6) +
  xlab("") +
  ylab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60))

ggplotly(weekly_team_track, tooltip = c("Vegas Win Projection %", 
                                        "Current Win %", 
                                        "Current Record",
                                        "Date")) %>%
  #  plotly::layout(margin = list(l = 30, r = 20, b = 5, t = 10))
  layout(
    margin = list(t = 120),
    title = list(
      text = paste("Team win percentage across season dates<br>",
                   "versus Vegas projection"),
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c("Current Win Percentage", rep("&nbsp;", 90), "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 100), "Date", "<br>"), collapse = "")
    ))
```


### By game number

```{r fig.height=10, fig.width=10}
team_track <- join_with_projections %>% 
  rename(`Over/Under \n Vegas \n Insider` = over_under) %>% 
  ggplot(aes(x = `Game Number`, y = `Current Win %`, 
             label = `Current Record`,
             text = `Game Result`, 
             color = `Over/Under \n Vegas \n Insider`)
  ) +
  geom_point() +
  geom_line() +
  geom_segment(aes(x = 1, xend = max(`Game Number`), 
                   y = `Vegas Win Projection %`, 
                   yend = `Vegas Win Projection %`),
               color = "blue") +
  facet_wrap(~ `Team Name`, nrow = 6) +
  xlab("") + #Game Number") +
  ylab("") + #<br><br>Current Win Percentage")
  theme_bw() +
  scale_color_manual(values = c("forestgreen", "black", "deeppink"))

ggplotly(team_track, tooltip = c("Vegas Win Projection %", 
                                 "Current Win %", 
                                 "Current Record",
                                 "Game Result")) %>%
  #  plotly::layout(margin = list(l = 30, r = 20, b = 5, t = 10))
  layout(
    margin = list(t = 120),
    title = list(
      text = paste("Team win percentage across games played <br>",
                   "versus Vegas projection"),
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c("Current Win Percentage", rep("&nbsp;", 100), "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 95), "Game Number", "<br>"), collapse = "")
    ))
```

### Approaching win projection {#win-proj}

- If a team exceeds the win projection during the season, 
**Wins Outcome Determined** will be shown as 
<span style="color:forestgreen;">OVER.</span>
- If a team is not able to win enough games to exceed the projection in the 
number of games remaining, **Wins Outcome Determined** will change to 
<span style="color:deeppink;">UNDER.</span>
- If either of these outcomes are still possible, **Wins Outcome Determined** 
will remain as Not Yet.

The purple square corresponds to the projected number of wins based on the current team winning percentage. The dotted purple line shows what this projection would look like from current outcomes.

```{r fig.height=10, fig.width=10}
options(scipen = 99)
team_wins_df <- join_with_projections %>% 
  mutate(
    `Over/Under Wins \n Vegas \n Insider` = current_wins > win_projection,
    `Wins Needed` = ceiling(win_projection) - current_wins,
    `Winning % In Remaining Games Needed` = round(
      `Wins Needed` / (num_games - current_wins - current_losses) * 100, 2),
    `Wins \n Outcome \n Determined` = case_when(
      `Winning % In Remaining Games Needed` <= 0 ~ "OVER",
      `Winning % In Remaining Games Needed` > 100 ~ "UNDER",
      TRUE ~ "Not Yet"
    )) %>% 
  rename(`Current Win Total` = current_wins)

current_team_wins_df <- team_wins_df %>% 
  group_by(`Team Name`) %>% 
  slice_tail(n = 1) %>% 
  arrange(`Wins Needed`) %>% 
  mutate(`Remaining Games` = num_games - `Game Number`) %>% 
  mutate(
    `Expected Wins At Current Pace` = round(
      `Remaining Games` * current_win_perc/100 + `Current Win Total`,
      2)
  ) %>% 
  mutate(spread = `Expected Wins At Current Pace` - win_projection) %>% 
  #  mutate(
  #    projected_slope = (`Expected Wins At Current Pace` - `Current Win Total`) / 
  #      (num_games - `Game Number`),
  #    projected_intercept = `Expected Wins At Current Pace` - num_games * projected_slope
  #    ) %>% 
  select(`Game Date`, `Team Name`, `Games Played` = `Game Number`,
         `Remaining Games`, `Current Winning %` = current_win_perc,
         `Current Win Total`,
         `Expected Wins At Current Pace`,
         `Vegas Win Projection` = win_projection,
         spread)#,
#         projected_slope,
#         projected_intercept)

current_team_wins_to_proj <- current_team_wins_df %>% 
  arrange(desc(spread)) %>% 
  pull(`Team Name`)

team_wins_track <- team_wins_df %>% 
  left_join(current_team_wins_df, 
            by = c("Game Date", "Team Name", "Current Win Total")) %>% 
  mutate(`Team Name` = factor(
    `Team Name`, 
    levels = current_team_wins_to_proj)
  ) %>% 
  ggplot(aes(label = `Current Record`,
             text = `Game Result`,
             color = `Wins \n Outcome \n Determined`)
  ) +
  geom_segment(aes(x = `Game Number`, xend = num_games,
                   y = `Current Win Total`,
                   yend = `Expected Wins At Current Pace`
  ),
  linetype = "dashed",
  color = "purple") +
  scale_x_continuous(limits = c(1, num_games)) + 
  geom_point(mapping = aes(x = `Game Number`, y = `Current Win Total`)) +
  geom_point(mapping = aes(x = num_games, y = `Expected Wins At Current Pace`), 
             color = "purple", shape = 15) +
  #  annotate("text", x = 4, y = 25, aes(label = `Wins Needed`)) +
  geom_segment(aes(x = 1, xend = num_games,
                   y = `Vegas Win Projection`,
                   yend = `Vegas Win Projection`),
               color = "blue") +
  facet_wrap(~ `Team Name`, nrow = 6) +
  xlab("") + #Game Number") +
  ylab("") + #<br><br>Current Win Percentage")
  theme_bw() +
  scale_color_manual(values = c("black", "forestgreen", "deeppink"))

ggplotly(team_wins_track, tooltip = c("Vegas Win Projection %", 
                                      "Expected Wins At Current Pace",
                                      "Current Win Total", 
                                      "Current Record",
                                      "Game Result")) %>%
  #  plotly::layout(margin = list(l = 30, r = 20, b = 5, t = 10))
  layout(
    margin = list(t = 120),
    title = list(
      text = paste("Team cumulative wins across games played <br>",
                   "versus Vegas wins projection"),
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c("Current Win Total", rep("&nbsp;", 60), "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 115), "Game Number", "<br>"), collapse = "")
    ))
```

***

#### Vegas Win Projections and Wins Needed

`Win Projection` was pulled from 
https://www.vegasinsider.com/nba/odds/win-totals/ prior to the season start. The last two
columns are based on games played through `r Sys.Date() - 1`.

```{r}
out_table |> 
  arrange(desc(`Wins To Go Over Vegas`), desc(`Remaining Games`)) |> 
  DT::datatable(
    rownames = FALSE,
    options = list(
      pageLength = 30, 
      dom = 'ft'
    ))
#  knitr::kable() %>% 
#  kableExtra::kable_styling("striped", full_width = F)
```





## Projected points for each player {.tabset .active}



#### Games through `r Sys.Date() - 1`

```{r}
# Custom CSS
custom_css <- tags$style(HTML("
  #proj_score table.dataTable tbody tr:nth-child(4) td {
    border-bottom: 3px solid black !important;
  }
"))

projected_score %>% 
  arrange(desc(Date), desc(`Projected Total Points`), 
          desc(`Number Correct`),
          desc(`Number (Wage 15) Correct`),
          desc(`Number (Wage 14) Correct`),
          desc(`Number (Wage 13) Correct`),
          desc(`Number (Wage 12) Correct`)) %>% 
  DT::datatable(
    options = list(
      pageLength = num_players,
      columnDefs = list(
        list(width = '80px', targets = 1)  # Increase the width of column index 1 (second column)
      )),
    elementId = "proj_score"
  ) %>%
  tagList(custom_css, .)
```

***


#### Point wagers by player with OVER(↑) / UNDER(↓) choice (highlighted by outcome determined)

```{r message=FALSE, warning=FALSE}
picks_wide <- picks %>% 
  unite("wage_choice", wage:choice, sep = " ") %>% 
  pivot_wider(names_from = player, values_from = wage_choice) %>% 
  inner_join(projections, by = "team") %>% 
  inner_join(standings, by = c("team" = "team_name")) %>% 
  mutate(percentage_projection = round(percentage_projection / 100, 3)) %>% 
  select(-conference, -division, -`Games Back`, -differential) %>% 
  mutate(loss_projection = num_games - win_projection, .after = win_projection) %>% 
  mutate(outcome_determined = case_when(
    wins > win_projection ~ "OVER",
    losses > loss_projection ~ "UNDER",
    TRUE ~ NA_character_)) %>% 
  mutate(team = case_when(
    outcome_determined == "OVER" ~ paste(team, knitr::asis_output("\U2191")),
    outcome_determined == "UNDER" ~ paste(team, knitr::asis_output("\U2193")),
    TRUE ~ team)) %>% 
  separate(
    col = Chester, 
    into = c("Chester_points", "chester_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Jake, 
    into = c("Jake_points", "jake_choice"), 
    sep = " "
  ) %>% 
  # separate(
  #   col = Jenelle, 
  #   into = c("Jenelle_points", "jenelle_choice"), 
  #   sep = " "
  # ) %>% 
  separate(
    col = Adonis, 
    into = c("Adonis_points", "adonis_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Mike, 
    into = c("Mike_points", "mike_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Mary, 
    into = c("Mary_points", "mary_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Ryan, 
    into = c("Ryan_points", "ryan_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Phil, 
    into = c("Phil_points", "phil_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Andy, 
    into = c("Andy_points", "andy_choice"), 
    sep = " "
  ) %>% 
  mutate(chester_correct = chester_choice == outcome_determined,
         jake_correct = jake_choice == outcome_determined,
         #         jenelle_correct = jenelle_choice == outcome_determined,
         adonis_correct = adonis_choice == outcome_determined,
         mike_correct = mike_choice == outcome_determined,
         mary_correct = mary_choice == outcome_determined,
         ryan_correct = ryan_choice == outcome_determined,
         phil_correct = phil_choice == outcome_determined,
         andy_correct = andy_choice == outcome_determined) %>% 
  mutate(Chester = if_else(
    chester_choice == "OVER",
    paste(Chester_points, knitr::asis_output("\U2191")),
    paste(Chester_points, knitr::asis_output("\U2193")))) %>% 
  mutate(Jake = if_else(
    jake_choice == "OVER",
    paste(Jake_points, knitr::asis_output("\U2191")),
    paste(Jake_points, knitr::asis_output("\U2193")))) %>% 
  # mutate(Jenelle = if_else(
  #   jenelle_choice == "OVER",
  #   paste(Jenelle_points, knitr::asis_output("\U2191")),
  #   paste(Jenelle_points, knitr::asis_output("\U2193")))) %>% 
  mutate(Adonis = if_else(
    adonis_choice == "OVER",
    paste(Adonis_points, knitr::asis_output("\U2191")),
    paste(Adonis_points, knitr::asis_output("\U2193")))) %>% 
  mutate(Mike = if_else(
    mike_choice == "OVER",
    paste(Mike_points, knitr::asis_output("\U2191")),
    paste(Mike_points, knitr::asis_output("\U2193")))) %>% 
  mutate(Mary = if_else(
    mary_choice == "OVER",
    paste(Mary_points, knitr::asis_output("\U2191")),
    paste(Mary_points, knitr::asis_output("\U2193")))) %>% 
  mutate(Ryan = if_else(
    ryan_choice == "OVER",
    paste(Ryan_points, knitr::asis_output("\U2191")),
    paste(Ryan_points, knitr::asis_output("\U2193")))) %>% 
  mutate(Phil = if_else(
    phil_choice == "OVER",
    paste(Phil_points, knitr::asis_output("\U2191")),
    paste(Phil_points, knitr::asis_output("\U2193")))) %>% 
  mutate(Andy = if_else(
    andy_choice == "OVER",
    paste(Andy_points, knitr::asis_output("\U2191")),
    paste(Andy_points, knitr::asis_output("\U2193")))) %>% 
  rename(`Target Wins` = win_projection) %>% 
  select(-abbreviation) %>% 
  select(Team = team, `Current Wins` = wins, `Target Wins`, 
         Adonis, Andy, Chester, Jake, #Jenelle, 
         Mary, Mike, Phil, Ryan, everything())
readr::write_rds(picks_wide %>% select(Team, Adonis_points:ryan_choice), "picks_wide.rds")
```

```{r message=FALSE, warning=FALSE}
cols_to_delete <- (num_players+3):(ncol(picks_wide) - 1)
do_not_show_list <-  as.list(cols_to_delete)

for (i in seq_along(do_not_show_list)) {
  names(do_not_show_list[[i]]) <- "targets"
}


for(i in (cols_to_delete - (num_players + 2)) ) {
  #  do_not_show_list[[i]]$targets <- as.list(cols_to_delete)[[i]]
  suppressWarnings(do_not_show_list[[i]]$visible <- FALSE)
}

#do_not_show_list[[33]] <- list(width = '180px', targets = 0)
do_not_show_list[[30]] <- list(width = '180px', targets = 0)

point_levels <- rep(seq(15, 6, by = -1), 2) %>% 
  sort(decreasing = TRUE)
paste_levels <- point_levels %>% 
  paste(c(knitr::asis_output("\U2191"), knitr::asis_output("\U2193")))

# sketch <- htmltools::withTags(table(
#   class = 'display',
#   thead(
#     tr(
#       th(colspan = 3, style = "border-right: solid 2px;", 'Target Wins'),
#       th(colspan = 2, 'Adonis')
#     ),
#     tr(
#       th("Team"),
#       th("Current Wins"),
#       th(style = "border-right: solid 2px;", "Target Wins")
#     )
#   )
# ))

picks_wide %>% 
  # Doesn't actually sort in DT
  mutate_at(
    .vars = c("Adonis", "Andy", "Chester", "Jake", #"Jenelle",
              "Mary", "Mike", "Phil", "Ryan"), 
    .funs = ~ factor(.x, levels = paste_levels)) %>% 
  #  purrr::map_df(rev) %>% 
  arrange(Team) %>% 
  DT::datatable(rownames = FALSE, options = list(
    pageLength = 30, dom = 'ft',
    autoWidth = TRUE, #scrollX = TRUE,
    columnDefs = do_not_show_list#,
    #   container = sketch
    #list(list(targets = as.list(cols_to_delete), visible = FALSE))
  )) %>% 
  formatStyle(
    "Chester", "chester_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Jake", "jake_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  # formatStyle(
  #   "Jenelle", "jenelle_correct",
  #   target = 'cell',
  #   backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
  #                                c("lightgreen", "hotpink", "white"))
  # ) %>% 
  formatStyle(
    "Adonis", "adonis_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Mike", "mike_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Mary", "mary_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Ryan", "ryan_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Phil", "phil_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Andy", "andy_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Team", target = 'cell', backgroundColor = "white"
  ) %>% 
  formatStyle(
    "Target Wins", target = 'cell', backgroundColor = "white", 
    `border-right` = "solid 2px"
  ) %>% 
  formatStyle(
    "Current Wins", target = 'cell', backgroundColor = "white"
  ) 
```

***

#### Point explanations by team for `r Sys.Date() - 1` {.tabset}

```{r message=FALSE}
library(DT)
player_table <- function(player_num) {
  df <- player_projections_by_team %>% 
    filter(Player == current_rankings[player_num],
           Date == Sys.Date() - 1) %>% 
    arrange(desc(`Current Projected Points`)) %>% 
    select(-Player, -Date)
  
  # if (sum(is.na(df$`Outcome Determined`)) == 30) {
  #   df %>% 
  #     select(-`Outcome Determined`) %>% 
  #     datatable(options = list(pageLength = 30, dom = 'ft'))
  # } else {
  df %>%
    datatable(rownames = FALSE,
              filter = 'top',
              extensions = "FixedColumns",
              options = list(pageLength = 30, dom = 'ft',
                             fixedColumns = list(leftColumns = 2))
    ) # %>%
  # formatStyle(
  #   "Outcome Determined",
  #   target = 'row',
  #   backgroundColor = styleEqual(c("OVER", "UNDER", NA_character_), 
  #                                c("lightgreen", "deeppink", "white"))
  # )
  # }
  
  # brks <- quantile(
  #   df$`Wins To Go Over Vegas`, 
  #   probs = seq(.05, .95, 0.1), 
  #   na.rm = TRUE
  # )
  # clrs <- rev(brewer.pal(n = 11, name = "PiYG"))
  # 
  # df %>% 
  #   datatable(options = list(pageLength = 30, dom = 'ft')) %>% 
  #   formatStyle(
  #     'Wins To Go Over Vegas',
  #     target = 'row',
  #     backgroundColor = styleInterval(brks, clrs)
  #   )
}
```


##### `r current_rankings[1]`

```{r}
player_table(1)
```

##### `r current_rankings[2]`

```{r}
player_table(2)
```

##### `r current_rankings[3]`

```{r}
player_table(3)
```

##### `r current_rankings[4]`

```{r}
player_table(4)
```

##### `r current_rankings[5]`

```{r}
player_table(5)
```

##### `r current_rankings[6]`

```{r}
player_table(6)
```

##### `r current_rankings[7]`

```{r}
player_table(7)
```

##### `r current_rankings[8]`

```{r}
player_table(8)
```

```{r, eval=FALSE}
player_table(9)
```


***

### All together

```{r fig.height=6, fig.width=10}
projected_score_highlight <- projected_score %>% 
  mutate(Player = factor(x = Player, levels = current_rankings)) %>%
  highlight_key(~Player)
projected_score_plot <- projected_score_highlight %>% 
  ggplot(aes(
    x = Date, y = `Projected Total Points`, color = Player)
  ) +
  geom_point() +
  geom_line() +
  xlab("") + 
  ylab("") + 
  scale_color_brewer(palette = "RdYlBu") +
  theme_bw()

gg <- ggplotly(projected_score_plot, 
               tooltip = c("Date", "Projected Total Points", "Player")) %>% 
  layout(
    margin = list(t = 80),
    title = list(
      text = "Projected points over time",
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c(rep("&nbsp;", 15), "Projected Total Points",  
          "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 9), "Date", "<br>"), collapse = "")
    ))
highlight(gg, color = c("lightgreen"), 
          off = "plotly_doubleclick",
          #          on = "plotly_hover", off = "plotly_deselect", 
          dynamic = FALSE)
```

### Small multiples

```{r fig.height=6, fig.width=10}
projected_score_facetted <- projected_score %>% 
  mutate(Player = factor(x = Player, levels = current_rankings)) %>% 
  ggplot(aes(
    x = Date, y = `Projected Total Points`)
  ) +
  geom_point() +
  geom_line() +
  xlab("") + 
  ylab("") + 
  theme_bw() +
  facet_wrap(~ Player, nrow = 2)

ggplotly(projected_score_facetted) %>% 
  layout(
    margin = list(t = 80),
    title = list(
      text = "Projected points (sorted from current best to worst) over time",
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c("Projected Total Points", rep("&nbsp;", 30), "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 105), "Date", "<br>"), collapse = "")
    ))
```



## Changes in projected points since yesterday

```{r results="asis"}
if (nrow(rows_today) == 0) {
  cat(paste("There was no change in scores for all players based on the",
            "outcomes of yesterday's games."))
}
```

```{r}
if (nrow(rows_today)) {
  changes_in_player_projections %>% 
    arrange(desc(Date), Team, Player) %>% 
    DT::datatable(options = list(pageLength = nrow(rows_today)))
}
```

```{r}
yesterday_scores <- scores %>% 
  filter(game_date == Sys.Date() - 1) %>% 
  select(`Game Date` = game_date,
         `Away Team` = away_team,
         `Away Score` = away_score,
         `Home Team` = home_team,
         `Home Score` = home_score)
```

***

```{r results="asis"}
if (nrow(yesterday_scores)) {
  cat("Yesterday's scores")
} else {
  cat("No NBA games were played yesterday.")
}
```

```{r}
if (nrow(yesterday_scores)) {
  yesterday_scores %>% knitr::kable() %>% 
    kableExtra::kable_styling("striped", full_width = F)
  #  DT::datatable(options = list(pageLength = nrow(yesterday_scores)))
}
```

```{r message=FALSE, warning=FALSE}
schedule <- read_csv(paste0("schedule-", this_year, "_initial.csv"))
                            #"-after-ist.csv"))
#                        ".csv")) #|> 
# rename(game_date = `Game Date`, start_time = `Start (ET)`,
#        away_team = `Visitor/Neutral`, home_team = `Home/Neutral`) |> 
# select(game_date, start_time, away_team, home_team) |> 
# mutate(game_date = lubridate::mdy(game_date))
```

```{r message=FALSE, warning=FALSE}
todays_games <- schedule %>% 
  filter(game_date == Sys.Date()) %>% # as.Date("2024-02-22")) |> 
  select(`Game Date` = game_date,
         `Start Time (Eastern)` = start_time,
         `Away Team` = away_team,
         `Home Team` = home_team)
```

***

```{r results="asis"}
if (nrow(todays_games)) {
  cat("Today's games")
} else {
  cat("No NBA games are scheduled for today.")
}
```

```{r}
if (nrow(todays_games)) {
  todays_games %>% knitr::kable() %>% 
    kableExtra::kable_styling("striped", full_width = F)
  #  DT::datatable(options = list(pageLength = nrow(yesterday_scores)))
}
```

```{r results="asis"}
if (nrow(todays_games)) {
  # Initialize a variable to track if any changes have been found
  any_changes_found <- FALSE
  
  # Function to simulate a game outcome and update standings
  simulate_game_outcome <- function(standings, winner, loser) {
    standings %>%
      mutate(wins = if_else(team_name == winner, wins + 1, wins),
             losses = if_else(team_name == loser, losses + 1, losses),
             `Winning Pct` = wins / (wins + losses))
  }
  
  # Total number of games in the season (for NBA, typically 82)
  num_games <- 82
  
  # Extract the current Over/Under status for each team based on the latest date
  current_over_under <- player_projections_by_team %>%
    arrange(Team, desc(Date)) %>%
    group_by(Team) %>%
    slice(1) %>%
    ungroup() %>%
    select(Team, `Current Over/Under`)
  
  # Extract the current clinched status for each team based on the latest date
  current_clinched_status <- player_projections_by_team |> 
    arrange(Team, desc(Date)) %>%
    group_by(Team) %>%
    slice(1) %>%
    ungroup() %>%
    select(Team, `Outcome Determined`)
  
  # Function to print potential changes in Over/Under status
  print_status_change <- function(simulated_standings, game_winner, current_over_under, projections) {
    changes <- simulated_standings %>%
      mutate(Projected_Wins = round(`Winning Pct` * num_games, 0)) %>%
      left_join(projections, by = c("team_name" = "team")) %>%
      mutate(
        Over_Under_Simulated = if_else(Projected_Wins > win_projection, "OVER", "UNDER"),
        Wins_Remaining = num_games - (wins + losses),
        Can_No_Longer_Win = wins + Wins_Remaining < win_projection,
        Can_No_Longer_Lose = losses + Wins_Remaining < num_games - win_projection
      ) %>%
      left_join(current_over_under, by = c("team_name" = "Team")) %>%
      left_join(current_clinched_status, by = c("team_name" = "Team")) |> 
      filter(((Over_Under_Simulated == "OVER" & `Current Over/Under` == "UNDER") |
                (Over_Under_Simulated == "UNDER" & `Current Over/Under` == "OVER") |
                Can_No_Longer_Win | Can_No_Longer_Lose) &
               (`Outcome Determined` == "not yet")) %>%
      select(team_name, `Current Over/Under`, Over_Under_Simulated, Can_No_Longer_Win, Can_No_Longer_Lose)
    
    if(nrow(changes) > 0) {
      any_changes_found <<- TRUE
      walk(seq_len(nrow(changes)), function(i) {
        row <- changes[i, ]
        if (row$Can_No_Longer_Win) {
          cat(glue("- If the {row$team_name} lose today's game, they clinch UNDER and can no longer reach the OVER projection."), "\n")
        } else if (row$Can_No_Longer_Lose) {
          cat(glue("- If the {row$team_name} win today's game, they clinch OVER and can no longer go UNDER based on losses."), "\n")
        } else if (game_winner == row$team_name) {
          cat(glue("- If the {game_winner} win today, they change from {row$`Current Over/Under`} to {row$Over_Under_Simulated}."), "\n")
        } else if (game_winner != row$team_name) {
          cat(glue("- If the {game_winner} win today, the {row$team_name} change from {row$`Current Over/Under`} to {row$Over_Under_Simulated}."), "\n")
        }
      })
    }
  }
  
  # Iterate over today's games and simulate outcomes
  for(i in 1:nrow(todays_games)) {
    away_team <- todays_games$`Away Team`[i]
    home_team <- todays_games$`Home Team`[i]
    
    # Simulate away team win and check for potential Over/Under status changes
    standings_away_win <- simulate_game_outcome(standings, away_team, home_team)
    print_status_change(standings_away_win, away_team, current_over_under, projections)
    
    # Simulate home team win and check for potential Over/Under status changes
    standings_home_win <- simulate_game_outcome(standings, home_team, away_team)
    print_status_change(standings_home_win, home_team, current_over_under, projections)
  }
  
  # After all simulations, check if any changes were found
  if(!any_changes_found) {
    cat("It is not possible for a change in Over/Under status based on the outcome of today's game(s).\n")
  }
}
```

## Current standings {.tabset}

### By conference {.tabset}

Note that tie breakers to determine playoff seeds are not currently set.

```{r eval=FALSE}
# 1. Prepare Winning Percentages
# 'standings' data frame has the winning percentages in the 'Winning Pct' column
winning_pct <- standings |>
  select(team_name, Winning_Pct = `Winning Pct`)

# 2. Filter Remaining Schedule
# Find the latest game date from 'scores_tidy'
latest_game_date <- max(scores_tidy$game_date)

# Filter the 'schedule' for games after the latest game date
remaining_schedule <- schedule |>
  filter(game_date > latest_game_date)

# 3. Calculate Remaining SOS for Each Team
# Combine the remaining schedule with the winning percentages of the opponents
remaining_schedule_with_wp <- remaining_schedule |>
  left_join(winning_pct, by = c("away_team" = "team_name")) |>
  rename(AwayTeamWP = Winning_Pct) |>
  left_join(winning_pct, by = c("home_team" = "team_name")) |>
  rename(HomeTeamWP = Winning_Pct)

# Calculate the total winning percentage of opponents for home games
home_sos <- remaining_schedule_with_wp %>%
  group_by(home_team) %>%
  summarise(Total_Opponent_WP = sum(AwayTeamWP, na.rm = TRUE),
            Total_Games = n())

# Calculate the total winning percentage of opponents for away games
away_sos <- remaining_schedule_with_wp %>%
  group_by(away_team) %>%
  summarise(Total_Opponent_WP = sum(HomeTeamWP, na.rm = TRUE),
            Total_Games = n())

# Combine the SOS for home and away
combined_sos <- bind_rows(home_sos %>% rename(team = home_team), 
                          away_sos %>% rename(team = away_team)) %>%
  group_by(team) %>%
  summarise(Total_Opponent_WP = sum(Total_Opponent_WP),
            Total_Games = sum(Total_Games)) %>%
  mutate(`Remaining Opponent Winning Percentage` = round(Total_Opponent_WP / Total_Games * 100, 3))

standings <- standings |> 
  inner_join(combined_sos |> select(-Total_Opponent_WP, -Total_Games), by = c("team_name" = "team"))
```

#### Western

```{r}
standings %>% 
  filter(conference == "West") %>% 
  select(-conference) %>% 
  relocate(division, .after = team_name) %>% 
  arrange(desc(differential)) %>% 
  select(-differential, -abbreviation) %>% 
  rename(
    Team = team_name, 
    Wins = wins, 
    Losses = losses, 
    Division = division
  ) %>% 
  knitr::kable(align = c('l', 'l', 'r', 'r', 'r', 'r')) %>% 
  kableExtra::kable_styling("striped", full_width = F)
```

#### Eastern

```{r}
standings %>% 
  filter(conference == "East") %>% 
  select(-conference) %>% 
  relocate(division, .after = team_name) %>% 
  arrange(desc(differential)) %>% 
  select(-differential, -abbreviation) %>% 
  rename(
    Team = team_name, 
    Wins = wins, 
    Losses = losses, 
    Division = division
  ) %>% 
  knitr::kable(align = c('l', 'l', 'r', 'r', 'r', 'r')) %>% 
  kableExtra::kable_styling("striped", full_width = F)
```


### By division {.tabset}

#### West {.tabset}

##### Pacific

```{r}
standings %>% 
  filter(division == "Pacific") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling("striped", full_width = F)
```

##### Northwest

```{r}
standings %>% 
  filter(division == "Northwest") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling("striped", full_width = F)
```

##### Southwest

```{r}
standings %>% 
  filter(division == "Southwest") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling("striped", full_width = F)
```

#### East {.tabset}

##### Central

```{r}
standings %>% 
  filter(division == "Central") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling("striped", full_width = F)
```

##### Atlantic

```{r}
standings %>% 
  filter(division == "Atlantic") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling("striped", full_width = F)
```

##### Southeast

```{r}
standings %>% 
  filter(division == "Southeast") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling("striped", full_width = F)
```
