---
title: "SOC 301-01: Social Statistics"
author: "Chester Ismay"
output:
  ioslides_presentation:
    incremental: no
    logo: ../figs/boxer.jpg
    fig.align: center
    widescreen: true
subtitle: Five Main Verbs + `%>%` (Practice)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center')
lc <- 0
library(ggplot2)
library(okcupiddata)
```

<style type="text/css">
    ol { list-style-type: upper-alpha; }
</style>

## Practice

We will use the `profiles` data set in the `okcupiddata` package and the `Titanic` (note the capital T) data set in the `datasets` package.  `datasets` comes automatically loaded with R.

```{r eval=FALSE}
library(dplyr)
library(okcupiddata)
data(profiles)
View(profiles)
Titanic <- data.frame(Titanic)
View(Titanic)
```


# Question 1

##

How many of those identified as `Male` survived the Titanic crash, broken down by Child/Adult?

>- `filter` to focus only on `Sex == "Male"`
>- `filter` to focus only on `Survived == "Yes"`
>- `group_by` to break down by `Age`
>- `summarize` to sum together the remaining counts of `Freq`

# Question 2

##

What's another `dplyr`-type question you can answer from the `Titanic` data set?

# Question 3

##

What is the mean `age` of OKCupid profiles corresponding to levels of `drugs` and `drinks`?

>- `group_by` to break down by `drugs` and `drinks`
>- `summarize` to calculate mean `age`

# Question 4

##

Produce an appropriate plot relating `age` to the levels of `drugs` and `drinks`.

# Question 5

##

What are the top 2 heights of those identifying as female entered based on `pets`?

>- `filter` to only choose `sex == "f"`
>- `select` to only choose `height` and `pets`
>- `group_by` to break down by `pets`
>- `top_n` to determine top 2 heights

# Question 6

##

Identify two new `dplyr` questions to answer based on the `profiles` data frame.

## Lab 6 next week

- Find a sociology related data set, download it to your computer, and upload it to the RStudio Server.  The best file formats are CSVs (comma-separated) files, but other file formats will work as well.
- Make a note about the "tidy"-ness of the data and how easy it is to understand the names of the variables.
- Find an article/data journalism piece related to sociology that uses data visualization to convey a message.  Write a two paragraph summary of the article and how data visualization is used.
- You can work in partners (2 per group) on this lab.

## To do for next time 
- Read over Chapter 6 in MODERN DIVE
- Read the article from Lifehacker published yesterday entitled [How to Lie to Yourself and Others With Statistics](http://lifehacker.com/how-to-lie-to-yourself-and-others-with-statistics-1788184031)
- We will go over the Learning Checks for Sections 6.1 and 6.2 on Monday.  Try them for practice before then.