---
title: "SOC 301-01: Social Statistics"
author: "Chester Ismay"
output:
  ioslides_presentation:
    incremental: no
    logo: ../figs/boxer.jpg
    fig.align: center
    widescreen: true
subtitle: Data Viz
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align = 'center')
```


## Learning Quote of the Day {.flexbox .vcenter}

> "It is not that I'm so smart. But I stay with the questions much longer."

― Albert Einstein

## Opening connections

On a half-sheet of paper, 

- Write your name on the front and answer
    - What are the differences between a histogram and a boxplot?
    - What is the difference between a histogram and a barplot?

- On the back of this sheet, answer the following questions:

    1. What is the definition of "observational unit"?
    2. What are the three properties of a **tidy** data set?
    3. What are the Five Named Graphs we explored in this chapter?


## 5NG and Grammar of Graphics Review

Run the following first in your console to create the data example:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Load packages
library(dplyr)
library(ggplot2)

# Create data frame
simple_ex <- data_frame(
    A = c(1, 2, 3, 4),
    B = c(1, 2, 3, 4),
    C = c(3, 2, 1, 2),
    D = c("a", "a", "b", "b")
  )
```

## View it

Let's view the data frame, which is in tidy format:

```{r, echo=TRUE, eval=FALSE}
View(simple_ex)
```

```{r, echo=FALSE}
simple_ex %>% knitr::kable()
```

## The Grammar of Graphics

* A statistical graphic is a `mapping` of `data` variables to `aes()`thetic attributes of `geom_`etric objects.
* A scatterplot has `point`s as the `geom_`etric object
* A linegraph has `line`s as the `geom_`etric object


## 1. Basic Scatterplot

* the `geom_`etric objects are `point`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`

```{r, fig.width=8, eval=FALSE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B)) + 
  geom_point()
```

## 1. Basic Scatterplot

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B)) + 
  geom_point()
```


## 2. Scatterplot with Color

* the `geom_`etric objects are `point`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`
    + `color` is variable `D`

```{r, eval=FALSE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B, color = D)) + 
  geom_point()
```


## 2. Scatterplot with Color

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B, color = D)) + 
  geom_point()
```



## 3. Scatterplot with Sizes

* the `geom_`etric objects are `point`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`
    + `size` is variable `C`

```{r, eval=FALSE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B, size = C)) + 
  geom_point()
```


## 3. Scatterplot with Sizes

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B, size = C)) + 
  geom_point()
```



## 4. Line Graph

* the `geom_`etric objects are `line`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`

```{r, eval=FALSE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B)) + 
  geom_line()
```


## 4. Line Graph

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B)) + 
  geom_line()
```


## 5. Line Graph with Color

* the `geom_`etric objects are `line`s
* the aesthetic attributes are:
    + `x`-axis is variable `A`
    + `y`-axis is variable `B`
    + `color` is variable `D`

```{r, eval=FALSE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B, color = D)) + 
  geom_line()
```

## 5. Line Graph with Color

```{r, fig.width=8, fig.height=4.5, echo=TRUE}
ggplot(data = simple_ex, mapping = aes(x = A, y = B, color = D)) + 
  geom_line()
```


# Finish Lab 3

## To do for next time

- Try the practice problems for the quiz [here](http://ismayc.github.io/teaching/soc301-f2016/slides/week-05/pq2all.html)
- No lab due on Tuesday next week

## Plan for next time

- Quiz #2 individually
- Quiz #2 in groups of 3 (of your choosing)
- Go over Quiz #2


    