---
title: "Simple Linear Regression Example"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
references:
- id: stat2
  title: STAT2 - Building Models for a World of Data
  author:
  - family: Cannon
    given: Ann R.
  - family: Cobb
    given: George W.
  - family: Hartlaub
    given: Bradley A.
  - family: Legler
    given: Julie M.
  - family: Lock
    given: Robin H.
  - family: Moore
    given: Thomas L.
  - family: Rossman
    given: Allan J.
  - family: Witmer
    given: Jeffrey A.
  type: book
  issued:
    year: 2013
---


```{r setup, include=FALSE}
pkg <- c("tidyr", "dplyr", "ggplot2", 
  "knitr", "rmarkdown", "readr", 
  "DT","devtools", "broom", "Stat2Data")

new.pkg <- pkg[!(pkg %in% installed.packages())]

if (length(new.pkg)) {
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
}

lapply(pkg, library, character.only = TRUE)

if(!require(oilabs))
  devtools::install_github("ismayc/oilabs", force = TRUE)

options(digits = 5, scipen = 99)
```

# Problem Statement

On January 28, 1986, a routine launch was anticipated for
the Challenger space shuttle. Seventy-three seconds into the 
flight, disaster happened: the shuttle
broke apart, killing all seven crew members on board. An investigation into the cause of the
disaster focused on a critical seal called an O-ring, and it is believed that damage to these O-rings
during a shuttle launch may be related to the ambient temperature during the launch. Observational data on O-rings for 23 randomly selected shuttle missions was collected, where the mission order
is based on the temperature at the time of the launch. `temp` gives the temperature in Fahrenheit and `damaged` is 1 when O-ring failed and 0 when it was undamaged. [Tweaked a bit from @oi [Chapter 8]]

# Competing Hypotheses

## In words

- Null hypothesis: The population coefficient on `temp` corresponding to the logistic regression modeling whether or not an O-ring failed as a function of temperature is zero.

- Alternative hypothesis:  The population coefficient on `temp` corresponding to the logistic regression modeling whether or not an O-ring failed as a function of temperature is nonzero.

## Another way with words

- Null hypothesis:  There is no association between temperature and whether or not an O-ring failed.

- Alternative hypothesis: There is an association between temperature and whether or not an O-ring failed.

## In symbols (with annotations)

- $H_0: \beta_1 = 0$, where $\beta_1$ represents the population coefficient on `temp` corresponding to the logistic regression modeling whether or not an O-ring failed as a function of temperature
- $H_A: \beta_1 \ne 0$

## Set $\alpha$

It's important to set the significance level before starting the testing using the data. Let's set the significance level at 5\% here.

# Exploring the sample data

```{r read_data}
library(dplyr)
library(knitr)
library(ggplot2)
library(oilabs)
orings <- read.delim("orings.txt")
```

The scatterplot below shows the relationship between `damage` and `temp`.  Note the use of `factor()` here which converts the `damage` variable into its two groups (1 and 0) instead of looking at it as a spectrum between 0 and 1.

```{r boxplot}
qplot(x = temp, y = factor(damage), 
  data = orings, geom = "point") + 
  geom_jitter(height = 0.4, alpha = 0.5)
```


## Guess about statistical significance

We are looking to see if `temp` is a significant predictor of `damage`.  Based on the plot, it seems reasonable to guess that we will reject the null hypothesis here.  There are no failures below 65 degrees and most of the points above 65 degrees did not fail.

# Check conditions

Remember that in order to use the shortcut (formula-based, theoretical) approach, we need to check that some conditions are met.

1. _Linear relationship between between linked response and predictors_:  
This condition is difficult to check without binning the data and computing something called the empirical logit function.  The plot of the data and the fitted logistic curve can usually provide a good idea on the quality of the fit.

    ```{r}
    qplot(x = temp, y = damage, 
        data = orings, geom = "point") +
      stat_smooth(method = "glm", 
        method.args = list(family = "binomial"),
        se = FALSE)
    ```

We have no clear reason to doubt the fit based on this plot.

2. _Independent observations_:  If cases are selected at random, the independent observations condition is met.

    The shuttle missions were selected at random here so the independent observations condition is met.


# Test statistic

The test statistic is a random variable based on the sample data. Here, we want to look at a way to estimate the population slope $\beta_1$.  A good guess is the sample coefficient $B_1$.  Recall that this sample coefficient in logistic regression is actually a random variable that will vary as different samples are (theoretically, would be) collected.  

We next look at our fitted curve from our sample of data:

```{r fit}
oring_glm <- glm(damage ~ temp, data = orings, family = binomial)
summary(oring_glm)
```

We are looking to see how likely is it for us to have observed a sample slope of $b_{1, obs} = -0.232$ or more extreme assuming that the population slope is 0 (assuming the null hypothesis is true).  If the conditions are met and assuming $H_0$ is true, we can "standardize" this original test statistic of $B_1$ into a $Z$ statistic that follows a $N(0,1)$:

$$ Z =\dfrac{ B_i - 0}{ {SE}_i } \sim N(0, 1) $$

where ${SE}_i$ represents the standard deviation of the distribution of the sample slopes.

## Observed test statistic

While one could compute this observed test statistic by "hand", the focus here is on the set-up of the problem and in understanding which formula for the test statistic applies.  We can use the `glm` function here to fit a curve and conduct the hypothesis test.  (We've already run this code earlier in the analysis, but it is shown here again for clarity.)

```{r}
oring_glm <- glm(damage ~ temp, data = orings, family = binomial)
summary(oring_glm)
```

We see here that the $b_{1, obs}$ value is around -0.232 which corresponds to a $z_{obs}$ value of -2.14.  Note that these values are negative.  This corresponds to a negative relationship between temperature and failure, which is what we saw in the scatterplot earlier.

# Compute $p$-value

The $p$-value---the probability of observing a $z}$ value of -2.14 or more extreme in our null distribution---is 0.032.  This can also be calculated in R directly:

```{r}
2 * pnorm(-2.14)
```

# State conclusion

We, therefore, have sufficient evidence to reject the null hypothesis.  Our initial guess that `temp` is a significant predictor of `damage` is supported by evidence.

---

