---
title: "Chi-Square - Test of Independence Example"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(infer)
library(janitor)
library(knitr)
options(digits = 5, scipen = 99)
```

# Problem Statement

A random sample of 500 U.S. adults were questioned regarding their political affiliation (`democrat` or `republican`) and opinion on a tax reform bill (`favor`, `indifferent`, `opposed`). Based on this sample, do we have reason to believe that political party and opinion on the bill are related? [Tweaked a bit from https://onlinecourses.science.psu.edu/stat500/node/56]

# Competing Hypotheses

## In words

- Null hypothesis: There is no association between the opinion on a tax reform bill and political affiliation for US adults.

- Alternative hypothesis:  There is an association between the opinion on a tax reform bill and political affiliation for US adults.

## Another way in words

- Null hypothesis: The long-run probability that a US adult who favors the bill identifies as Democrat is the **same** as the long-run probability that a US adult who is indifferent towards the bill identifies as Democrat, which is the **same** as the long-run probability that a US adult who is opposed to the bill identifies as Democrat. In other words all three long-run probabilities are actually the same.

- Alternative hypothesis:  At least one of these parameter (long-run) probabilities is different from the others

## In symbols (with annotations)

- $H_0: P_{favor} = P_{indifferent} =  P_{opposed}$, where $P$ represents the long-run probability of being a Democrat.
- $H_A$: At least one of these parameter probabilities is different from the others

## Set $\alpha$

It's important to set the significance level before starting the testing using the data. Let's set the significance level at 5\% here.

# Exploring the sample data

```{r read_data}
tax_opinion <- read_csv("party_tax.csv")
```

```{r summarize}
tax_opinion |>
  tabyl(opinion, party)
```

```{r stacked_bar}
ggplot(data = tax_opinion, mapping = aes(x = party, fill = opinion)) +
  geom_bar()
```

## Guess about statistical significance

We are looking to see if a difference exists in the distribution of opinions across the two parties.  Based on the bar plot, we have reason to believe that a difference exists since the `favor`/`opposed` split looks different between Democrats and Republicans, BUT...it's important to use statistics to see if that difference is actually statistically significant!

***

# Non-traditional methods

We'll use the **`infer`** package to carry out the simulation-based workflow with the $\chi^2$ statistic.

## Observed statistic

```{r obs_stat}
chisq_hat <- tax_opinion |>
  specify(opinion ~ party) |>
  hypothesize(null = "independence") |>
  calculate(stat = "Chisq")
chisq_hat
```

## Randomization for Hypothesis Test

Under the null of no association, the `opinion` and `party` labels are exchangeable, so `infer` shuffles them and recomputes the $\chi^2$ statistic 10,000 times to build the null distribution.

```{r sim}
set.seed(2018)
null_distn <- tax_opinion |>
  specify(opinion ~ party) |>
  hypothesize(null = "independence") |>
  generate(reps = 10000, type = "permute") |>
  calculate(stat = "Chisq")
```

The $\chi^2$ statistic is always non-negative, so this is a one-sided (right-tailed) test.

```{r shade_pvalue}
null_distn |>
  visualize() +
  shade_p_value(obs_stat = chisq_hat, direction = "greater")
```

We can also overlay the theoretical $\chi^2$ distribution:

```{r overlay}
null_distn |>
  visualize(method = "both") +
  shade_p_value(obs_stat = chisq_hat, direction = "greater")
```

### Calculate $p$-value

```{r get_pvalue}
pvalue <- null_distn |>
  get_p_value(obs_stat = chisq_hat, direction = "greater")
pvalue
```

So our $p$-value is `r pvalue$p_value` and we reject the null hypothesis at the 5% level.

***

# Traditional methods

## Check conditions

Remember that in order to use the short-cut (formula-based, theoretical) approach, we need to check that some conditions are met.

1. _Independence_:  Each case that contributes a count to the table must be independent of all the other cases in the table.

    This condition is met since cases were selected at random to observe.

2. _Sample size_:  Each cell count must have at least 5 expected cases.

    This is met by observing the table above.

3. _Degrees of freedom_:  We need 3 or more columns (or rows) in the table.

    This is met by observing the table above.

## Test statistic

The test statistic measures deviations from what we would expect in the cells of the table if the null hypothesis were true.  This requires us to calculate expected counts via

$$\text{Expected Count}_{\text{row } i, \text{col } j} = \dfrac{\text{row } i \text{ total} \times \text{column } j \text{ total}}{\text{table total}}$$

$$X^2 = \sum_{\text{all cells in the table}} \dfrac{(\text{observed count} - \text{expected count})^2}{\text{expected count}}$$

Assuming the conditions outlined above are met, $X^2 \sim \chi^2(df = (R - 1) \times (C - 1))$ where $R$ is the number of rows in the table and $C$ is the number of columns.

## Observed test statistic

We can use `infer`'s `chisq_test()` wrapper to perform the theory-based test for us, which reports the $\chi^2$ statistic, degrees of freedom, and $p$-value.

```{r chisq_test}
tax_opinion |>
  chisq_test(opinion ~ party)
```

## Compute $p$-value

The theory-based $p$-value closely matches the simulation-based $p$-value from above, both leading to the same conclusion.

# State conclusion

We, therefore, have sufficient evidence to reject the null hypothesis.  Our initial guess that a statistically significant difference existed in the proportions of Democrats across the three groups was backed up by this statistical analysis.  We do have evidence to suggest that there is a dependency between the position taken on the tax reform bill and political party for US adults, based on this sample.

***

# Comparing results

The simulation-based and theory-based approaches give very similar $p$-values here, as the overlaid distributions above suggest.  Since the conditions for the theory-based test are met, both the randomization and $\chi^2$-distribution approaches lead to the same conclusion.
