---
title: "Two Proportions Example"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
references:
- id: oi
  title: OpenIntro Statistics, Third Edition
  author:
  - family: Diez
    given: David
  - family: Barr
    given: Christopher
  - family: Cetinkaya-Rundel
    given: Mine
  type: book
  issued:
    year: 2015
---


```{r setup, include=FALSE}
pkg <- c("tidyr", "dplyr", "ggplot2", 
  "knitr", "rmarkdown", "readr", 
  "DT","devtools", "broom")

new.pkg <- pkg[!(pkg %in% installed.packages())]

if (length(new.pkg)) {
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
}

lapply(pkg, library, character.only = TRUE)

if(!require(oilabs))
  devtools::install_github("ismayc/oilabs")

options(digits = 5, scipen = 99)
```

# Problem Statement

A 2010 survey asked 827 randomly sampled registered voters
in California "Do you support? Or do you oppose? Drilling for oil and natural gas off the Coast of
California? Or do you not know enough to say?" Conduct a hypothesis test to determine if the data
provide strong evidence that the proportion of college
graduates who do not have an opinion on this issue is
different than that of non-college graduates. [Tweaked a bit from @oi [Chapter 6]]

# Competing Hypotheses

## In words

- Null hypothesis: There is no association between having an opinion on drilling and having a college degree for all registered California voters in 2010.

- Alternative hypothesis:  There is an association between having an opinion on drilling and having a college degree for all registered California voters in 2010.

## Another way in words

- Null hypothesis: The probability that a Californian voter in 2010 having no opinion on drilling and is a college graduate is the **same** as that of a non-college graduate.

- Alternative hypothesis:  These parameter probabilities are different.

## In symbols (with annotations)

- $H_0: P_{college} = P_{no\_college}$ or $H_0: P_{college} - P_{no\_college} = 0$, where $P$ represents the probability of not having an opinion on drilling.
- $H_A: P_{college} - P_{no\_college} \ne 0$

## Set $\alpha$

It's important to set the significance level before starting the testing using the data. Let's set the significance level at 5\% here.

# Exploring the sample data

```{r read_data, echo=FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
library(oilabs)
#download.file("https://raw.githubusercontent.com/ismayc/ismayc.github.io/master/teaching/sample_problems/offshore_drilling.txt", destfile = "offshore_drilling.txt")
offshore <- read.delim("offshore_drilling.txt") %>%
  mutate(response = ifelse(position == "do not know", "no opinion", "opinion")) %>%
  select(-position)
str(offshore)
```

```{r summarize}
table(offshore$college_grad, offshore$response)
```

```{r mosaic}
mosaicplot(table(offshore$college_grad, offshore$response),
           xlab = "College Grad", 
           ylab = "Opinion",
           main = "College Grad vs Opinion",
           color = c("salmon", "blue"))
```

```{r stacked_bar}
qplot(x = college_grad, data = offshore, fill = response, geom = "bar")
```

## Guess about statistical significance

We are looking to see if a difference exists in the heights of the bars corresponding to `no opinion` for the mosaic plot.  Based solely on mosaicplot, we have little reason to believe that a difference exists since the bars seem to be about the same height, BUT...it's important to use statistics to see if that difference is actually statistically significant!

# Check conditions

Remember that in order to use the short-cut (formula-based, theoretical) approach, we need to check that some conditions are met.

1. _Independent observations_:  Each case that was selected must be independent of all the other cases selected.

    This condition is met since cases were selected at random to observe.

2. _Sample size_:  The number of pooled successes and pooled failures must be at least 10 for each group.

    We need to first figure out the pooled success rate:  $$\hat{p}_{obs} = \dfrac{131 + 104}{827} = 0.28.$$  We now determine expected (pooled) success and failure counts:
    
    $0.28 \cdot (131 + 258) = `r 0.28 * (131 + 258)`$, $0.72 \cdot (131 + 258) = `r 0.72 * (131 + 258)`$
    
    $0.28 \cdot (104 + 334) = `r 0.28 * (104 + 334)`$, $0.72 \cdot (104 + 334) = `r 0.72 * (104 + 334)`$

3. _Independent selection of samples_:  The cases are not paired in any meaningful way.

    We have no reason to suspect that a college graduate selected would have any relationship to a non-college graduate selected.

# Test statistic

The test statistic is a random variable based on the sample data.  Here, we are interested in seeing if our observed difference in sample proportions corresponding to no opinion on drilling ($\hat{p}_{college, obs} - \hat{p}_{no\_college, obs}$ = `r prop.table(table(offshore$college_grad, offshore$response))[1, 1] - prop.table(table(offshore$college_grad, offshore$response))[2, 1]`) is statistically different than 0.  Assuming that conditions are met and the null hypothesis is true, we can use the standard normal distribution to standardize the difference in sample proportions ($\hat{P}_{college} - \hat{P}_{no\_college}$) using the standard error of $\hat{P}_{college} - \hat{P}_{no\_college}$ and the pooled estimate:

$$ Z =\dfrac{ (\hat{P}_1 - \hat{P}_2) - 0}{\sqrt{\dfrac{\hat{P}(1 - \hat{P})}{n_1} + \dfrac{\hat{P}(1 - \hat{P})}{n_2} }} \sim N(0, 1) $$ where $\hat{P} = \dfrac{\text{total number of successes} }{ \text{total number of cases}}.$

## Observed test statistic

While one could compute this observed test statistic by "hand", the focus here is on the set-up of the problem and in understanding which formula for the test statistic applies.  We can use the `inference` function in the `oilabs` package to perform this analysis for us.

```{r infer}
inference(x = offshore$college_grad, 
          y = offshore$response, 
          est = "proportion", 
          success = "no opinion",
          alternative = "twosided", 
          type = "ht", 
          null = 0,
          method = "theoretical", 
          eda_plot = FALSE, 
          inf_plot = FALSE)
```

We see here that the $z_{obs}$ value is 3.161.  The observed difference in sample proportions is 3.161 standard deviations larger than 0.

# Compute $p$-value

The $p$-value---the probability of observing a $Z$ value of 3.161 or more extreme in our null distribution---is 0.0015.  This can also be calculated in R directly:

```{r pval}
2 * pnorm(3.161, lower.tail = FALSE)
```

Note that we could also do this test directly without invoking the `inference` function using the `prop.test` function.

```{r prop.test}
prop.test(x = table(offshore$college_grad, offshore$response),
       n = nrow(offshore),
       alternative = "two.sided",
       correct = FALSE)
```

`prop.test` does a $\chi^2$ test here but this matches up exactly with what we would expect:  $x^2_{obs} = 9.99 = (3.16)^2 = (z_{obs})^2$ and the $p$-values are the same because we are focusing on a two-tailed test.


# State conclusion

We, therefore, have sufficient evidence to reject the null hypothesis.  Our initial guess that a statistically significant difference did not exist in the proportions of no opinion on offshore drilling between college educated and non-college educated Californians was not validated.  We do have evidence to suggest that there is a dependency between college graduation and position on offshore drilling for Californians.

---

