---
title: "Two Means - Paired Samples Example"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
pkg <- c("tidyr", "dplyr", "ggplot2", 
  "knitr", "rmarkdown", "readr", "mosaic",
  "DT","devtools", "broom", "tibble")

new.pkg <- pkg[!(pkg %in% installed.packages())]

if (length(new.pkg)) {
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
}

lapply(pkg, library, character.only = TRUE)

#if(!require(oilabs))
#  devtools::install_github("ismayc/oilabs", force = TRUE)

options(digits = 2, scipen = 99)
```


# Problem Statement

Trace metals in drinking water affect the flavor and an unusually high concentration can pose a health hazard. Ten pairs of data were taken measuring zinc concentration in bottom water and surface water at 10 randomly selected locations on a stretch of river. Do the data suggest that the true average concentration in the bottom water exceeds that of surface water? [Tweaked a bit from https://onlinecourses.science.psu.edu/stat500/node/51]

# Competing Hypotheses

## In words

- Null hypothesis: The mean concentration in the bottom water is the same as that of the surface water at different paired locations.

- Alternative hypothesis:  The mean concentration in the bottom water is greater than that of the surface water at different paired locations.

## In symbols (with annotations)

- $H_0: \mu_{diff} = 0$, where $\mu_{diff}$ represents the mean difference in concentration for bottom water minus surface water.
- $H_A: \mu_{diff} > 0$

## Set $\alpha$

It's important to set the significance level before starting the testing using the data. Let's set the significance level at 5% here.

# Exploring the sample data

```{r read_data}
library(dplyr)
library(knitr)
library(ggplot2)
library(oilabs)
download.file("https://raw.githubusercontent.com/ismayc/ismayc.github.io/master/teaching/sample_problems/zinc_conc.txt", destfile = "zinc_conc.txt")
zinc_conc <- read.delim("zinc_conc.txt") %>%
  mutate(diff = bottom - surface)
```

```{r summarize}
zinc_summ <- zinc_conc %>%
  summarize(sample_size = n(),
    mean = mean(diff),
    sd = sd(diff),
    minimum = min(diff),
    lower_quartile = quantile(diff, 0.25),
    median = median(diff),
    upper_quartile = quantile(diff, 0.75),
    max = max(diff))
kable(zinc_summ)
```

The histogram below also shows the distribution of `diff`.

```{r hist1}
qplot(x = diff, data = zinc_conc, geom = "histogram", 
      binwidth = 0.04, col = I("white"))
```


## Guess about statistical significance

We are looking to see if the sample mean of `r zinc_summ$mean` is statistically greater than $\mu_0 = 23$.  They seem to be quite close, but we have a small number of pairs here.  Let's guess that we will fail to reject the null hypothesis. 

# Check conditions

Remember that in order to use the shortcut (formula-based, theoretical) approach, we need to check that some conditions are met.

1. _Independent observations_:  The observations among pairs are independent.

   The locations are selected independently through random sampling so this condition is met.

2. _Approximately normal_:  The distribution of population of differences is normal or the number of pairs is at least 30.

    The histogram above does show some skew so we have reason to doubt the population being normal based on this sample.  The Q-Q plot below also shows some skew.  The small number of pairs may be a problem here.

```{r qqplot}
qplot(sample = diff, data = zinc_conc)
```

# Test statistic

The test statistic is a random variable based on the sample data.  Here, we want to look at a way to estimate the population mean difference $\mu_{diff}$.  A good guess is the sample mean difference $\bar{X}_{diff}$.  Recall that this sample mean is actually a random variable that will vary as different samples are (theoretically, would be) collected.  We are looking to see how likely is it for us to have observed a sample mean of $\bar{x}_{diff, obs} = 0.0804$ or larger assuming that the population mean difference is 0 (assuming the null hypothesis is true).  If the conditions are met and assuming $H_0$ is true, we can "standardize" this original test statistic of $\bar{X}_{diff}$ into a $T$ statistic that follows a $t$ distribution with degrees of freedom equal to $df = n - 1$:

$$ T =\dfrac{ \bar{X}_{diff} - 0}{ S_{diff} / \sqrt{n} } \sim t (df = n - 1) $$

where $S$ represents the standard deviation of the sample differences and $n$ is the number of pairs.

## Observed test statistic

While one could compute this observed test statistic by "hand", the focus here is on the set-up of the problem and in understanding which formula for the test statistic applies.  We can use the `inference` function in the `oilabs` package to perform this analysis for us. 

```{r inference}
inference(y = zinc_conc$diff, 
          est = "mean", 
          null = 0, 
          alternative = "greater",
          type = "ht", 
          method = "theoretical", 
          eda_plot = FALSE, 
          inf_plot = FALSE)
```

We see here that the $t_{obs}$ value is around 4.864.

# Compute $p$-value

The $p$-value---the probability of observing an $t_{obs}$ value of 4.864 or more in our null distribution of a $t$ with 9 degrees of freedom---is 0.0004.  This can also be calculated in R directly:

```{r pval1}
pt(4.864, df = nrow(zinc_conc) - 1, lower.tail = FALSE)
```

Note that we could also do this test directly without invoking the `inference` function using the `t.test` function.

```{r t.test}
t.test(x = zinc_conc$diff, 
       alternative = "greater",
       mu = 0)
```

# State conclusion

We, therefore, have sufficient evidence to reject the null hypothesis.  Our initial guess that our observed sample mean difference was not statistically greater than the hypothesized mean of 0 has been invalidated here.  Based on this sample, we have evidence that the mean concentration in the bottom water is greater than that of the surface water at different paired locations.

---

