---
title: 'NBA Over/Under 2020-2021 Season'
author: Created by Chester Ismay
output: html_document
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path('../2021-nba-over-under.html')) })
---

```{r include=FALSE}
source("01-03.R")
knitr::opts_chunk$set(echo = FALSE)
library(lubridate)
library(DT)
```

Last updated at `r Sys.time()` US Pacific time

# {.tabset}

## By team {.tabset}

### By week

This plot is updated on Monday each week with the previous week's results.

```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
weekly_standings <- standings_grid %>% 
  mutate(day_of_week = wday(date, label = TRUE)) %>% 
  filter(day_of_week == "Mon") %>% 
  select(-day_of_week) %>% 
  rename(Date = date, `Team Name` = team)

weekly_team_track <- ggplot(
  weekly_standings,
  aes(x = Date, y = `Current Win %`, label = `Current Record`)
) +
  geom_point() +
  geom_line() +
  geom_segment(aes(x = min(Date), xend = max(Date), 
                   y = `Vegas Insider Win Projection %`, 
                   yend = `Vegas Insider Win Projection %`),
               color = "blue") +
  facet_wrap(~ `Team Name`, nrow = 6) +
  xlab("") +
  ylab("") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60))

ggplotly(weekly_team_track, tooltip = c("Vegas Insider Win Projection %", 
                                        "Current Win %", 
                                        "Current Record",
                                        "Date")) %>%
  #  plotly::layout(margin = list(l = 30, r = 20, b = 5, t = 10))
  layout(
    margin = list(t = 120),
    title = list(
      text = paste("Team win percentage across season dates<br>",
                   "versus Vegas Insider projection"),
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c("Current Win Percentage", rep("&nbsp;", 90), "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 100), "Date", "<br>"), collapse = "")
    ))
```


### By game number

```{r fig.height=10, fig.width=10}
team_track <- join_with_projections %>% 
  rename(`Over/Under \n Vegas \n Insider` = over_under) %>% 
  ggplot(aes(x = `Game Number`, y = `Current Win %`, 
             label = `Current Record`,
             text = `Game Result`, 
             color = `Over/Under \n Vegas \n Insider`)
  ) +
  geom_point() +
  geom_line() +
  geom_segment(aes(x = 1, xend = max(`Game Number`), 
                   y = `Vegas Insider Win Projection %`, 
                   yend = `Vegas Insider Win Projection %`),
               color = "blue") +
  facet_wrap(~ `Team Name`, nrow = 6) +
  xlab("") + #Game Number") +
  ylab("") + #<br><br>Current Win Percentage")
  theme_bw() +
  scale_color_manual(values = c("forestgreen", "black", "deeppink"))

ggplotly(team_track, tooltip = c("Vegas Insider Win Projection %", 
                                 "Current Win %", 
                                 "Current Record",
                                 "Game Result")) %>%
  #  plotly::layout(margin = list(l = 30, r = 20, b = 5, t = 10))
  layout(
    margin = list(t = 120),
    title = list(
      text = paste("Team win percentage across games played <br>",
                   "versus Vegas Insider projection"),
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c("Current Win Percentage", rep("&nbsp;", 90), "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 115), "Game Number", "<br>"), collapse = "")
    ))
```

### Approaching win projection {#win-proj}

- If a team exceeds the win projection during the season, 
**Wins Outcome Determined** will be shown as 
<span style="color:forestgreen;">OVER.</span>
- If a team is not able to win enough games to exceed the projection in the 
number of games remaining, **Wins Outcome Determined** will change to 
<span style="color:deeppink;">UNDER.</span>
- If either of these outcomes are still possible, **Wins Outcome Determined** 
will remain as Not Yet.

The purple square corresponds to the projected number of wins based on the current team winning percentage. The dotted purple line shows what this projection would look like from current outcomes.

```{r fig.height=10, fig.width=10}
options(scipen = 99)
team_wins_df <- join_with_projections %>% 
  mutate(
    `Over/Under Wins \n Vegas \n Insider` = current_wins > win_projection,
    `Wins Needed` = ceiling(win_projection) - current_wins,
    `Winning % In Remaining Games Needed` = round(
      `Wins Needed` / (72 - current_wins - current_losses) * 100, 2),
    `Wins \n Outcome \n Determined` = case_when(
      `Winning % In Remaining Games Needed` <= 0 ~ "OVER",
      `Winning % In Remaining Games Needed` > 100 ~ "UNDER",
      TRUE ~ "Not Yet"
    )) %>% 
  rename(`Current Win Total` = current_wins)

current_team_wins_df <- team_wins_df %>% 
  group_by(`Team Name`) %>% 
  slice_tail(n = 1) %>% 
  arrange(`Wins Needed`) %>% 
  mutate(`Remaining Games` = 72 - `Game Number`) %>% 
  mutate(
    `Expected Wins At Current Pace` = round(
      `Remaining Games` * current_win_perc/100 + `Current Win Total`,
      2)
  ) %>% 
  mutate(spread = `Expected Wins At Current Pace` - win_projection) %>% 
  #  mutate(
  #    projected_slope = (`Expected Wins At Current Pace` - `Current Win Total`) / 
  #      (72 - `Game Number`),
  #    projected_intercept = `Expected Wins At Current Pace` - 72 * projected_slope
  #    ) %>% 
  select(`Game Date`, `Team Name`, `Games Played` = `Game Number`,
         `Remaining Games`, `Current Winning %` = current_win_perc,
         `Current Win Total`,
         `Expected Wins At Current Pace`,
         `Vegas Insider Win Projection` = win_projection,
         spread)#,
#         projected_slope,
#         projected_intercept)

current_team_wins_to_proj <- current_team_wins_df %>% 
  arrange(desc(spread)) %>% 
  pull(`Team Name`)

team_wins_track <- team_wins_df %>% 
  left_join(current_team_wins_df, 
            by = c("Game Date", "Team Name", "Current Win Total")) %>% 
  mutate(`Team Name` = factor(
    `Team Name`, 
    levels = current_team_wins_to_proj)
  ) %>% 
  ggplot(aes(label = `Current Record`,
             text = `Game Result`,
             color = `Wins \n Outcome \n Determined`)
  ) +
  geom_segment(aes(x = `Game Number`, xend = 72,
                   y = `Current Win Total`,
                   yend = `Expected Wins At Current Pace`
  ),
  linetype = "dashed",
  color = "purple") +
  scale_x_continuous(limits = c(1, 72)) + 
  geom_point(mapping = aes(x = `Game Number`, y = `Current Win Total`)) +
  geom_point(mapping = aes(x = 72, y = `Expected Wins At Current Pace`), 
             color = "purple", shape = 15) +
  #  annotate("text", x = 4, y = 25, aes(label = `Wins Needed`)) +
  geom_segment(aes(x = 1, xend = 72,
                   y = `Vegas Insider Win Projection`,
                   yend = `Vegas Insider Win Projection`),
               color = "blue") +
  facet_wrap(~ `Team Name`, nrow = 6) +
  xlab("") + #Game Number") +
  ylab("") + #<br><br>Current Win Percentage")
  theme_bw() +
  scale_color_manual(values = c("black", "forestgreen", "deeppink"))

ggplotly(team_wins_track, tooltip = c("Vegas Insider Win Projection %", 
                                      "Expected Wins At Current Pace",
                                      "Current Win Total", 
                                      "Current Record",
                                      "Game Result")) %>%
  #  plotly::layout(margin = list(l = 30, r = 20, b = 5, t = 10))
  layout(
    margin = list(t = 120),
    title = list(
      text = paste("Team cumulative wins across games played <br>",
                   "versus Vegas Insider wins projection"),
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c("Current Win Total", rep("&nbsp;", 60), "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 115), "Game Number", "<br>"), collapse = "")
    ))
```

***

#### Vegas Insider Win Projections

These projections were pulled from https://www.vegasinsider.com/nba/odds/win-totals/
on 2020-11-30.

```{r}
projections %>% 
  select(Team = team,
         `Win Projection` = win_projection) %>% 
  knitr::kable()
```





## Projected points for each player {.tabset .active}



#### Games through `r Sys.Date() - 1`

```{r}
projected_score %>% 
  arrange(desc(Date), desc(`Projected Total Points`)) %>% 
  DT::datatable(options = list(pageLength = 8))
```

***


#### Point wagers by player with OVER(↑) / UNDER(↓) choice (highlighted by outcome determined)

```{r message=FALSE, warning=FALSE}
picks_wide <- picks %>% 
  unite("wage_choice", wage:choice, sep = " ") %>% 
  pivot_wider(names_from = player, values_from = wage_choice) %>% 
  inner_join(projections, by = "team") %>% 
  inner_join(standings, by = c("team" = "team_name")) %>% 
  mutate(percentage_projection = round(percentage_projection / 100, 3)) %>% 
  select(-conference, -division, -`Games Back`, -differential) %>% 
  mutate(loss_projection = 72 - win_projection, .after = win_projection) %>% 
  mutate(outcome_determined = case_when(
    wins > win_projection ~ "OVER",
    losses > loss_projection ~ "UNDER",
    TRUE ~ NA_character_)) %>% 
  mutate(team = case_when(
    outcome_determined == "OVER" ~ paste(team, knitr::asis_output("\U2193")),
    outcome_determined == "UNDER" ~ paste(team, knitr::asis_output("\U2191")),
    TRUE ~ team)) %>% 
  separate(
    col = Chester, 
    into = c("Chester_points", "chester_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Jake, 
    into = c("Jake_points", "jake_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Jenelle, 
    into = c("Jenelle_points", "jenelle_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Adonis, 
    into = c("Adonis_points", "adonis_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Mike, 
    into = c("Mike_points", "mike_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Mary, 
    into = c("Mary_points", "mary_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Ryan, 
    into = c("Ryan_points", "ryan_choice"), 
    sep = " "
  ) %>% 
  separate(
    col = Phil, 
    into = c("Phil_points", "phil_choice"), 
    sep = " "
  ) %>% 
  mutate(chester_correct = chester_choice == outcome_determined,
         jake_correct = jake_choice == outcome_determined,
         jenelle_correct = jenelle_choice == outcome_determined,
         adonis_correct = adonis_choice == outcome_determined,
         mike_correct = mike_choice == outcome_determined,
         mary_correct = mary_choice == outcome_determined,
         ryan_correct = ryan_choice == outcome_determined,
         phil_correct = phil_choice == outcome_determined) %>% 
  mutate(Chester = if_else(
    chester_choice == "OVER",
    paste(Chester_points, knitr::asis_output("\U2193")),
    paste(Chester_points, knitr::asis_output("\U2191")))) %>% 
  mutate(Jake = if_else(
    jake_choice == "OVER",
    paste(Jake_points, knitr::asis_output("\U2193")),
    paste(Jake_points, knitr::asis_output("\U2191")))) %>% 
  mutate(Jenelle = if_else(
    jenelle_choice == "OVER",
    paste(Jenelle_points, knitr::asis_output("\U2193")),
    paste(Jenelle_points, knitr::asis_output("\U2191")))) %>% 
  mutate(Adonis = if_else(
    adonis_choice == "OVER",
    paste(Adonis_points, knitr::asis_output("\U2193")),
    paste(Adonis_points, knitr::asis_output("\U2191")))) %>% 
  mutate(Mike = if_else(
    mike_choice == "OVER",
    paste(Mike_points, knitr::asis_output("\U2193")),
    paste(Mike_points, knitr::asis_output("\U2191")))) %>% 
  mutate(Mary = if_else(
    mary_choice == "OVER",
    paste(Mary_points, knitr::asis_output("\U2193")),
    paste(Mary_points, knitr::asis_output("\U2191")))) %>% 
  mutate(Ryan = if_else(
    ryan_choice == "OVER",
    paste(Ryan_points, knitr::asis_output("\U2193")),
    paste(Ryan_points, knitr::asis_output("\U2191")))) %>% 
  mutate(Phil = if_else(
    phil_choice == "OVER",
    paste(Phil_points, knitr::asis_output("\U2193")),
    paste(Phil_points, knitr::asis_output("\U2191")))) %>% 
  rename(`Win Target` = win_projection) %>% 
  select(Team = team, `Current Wins` = wins, `Win Target`, 
         Adonis, Chester, Jake, Jenelle, Mary, Mike, Phil, Ryan, 
         everything())
```

```{r message=FALSE, warning=FALSE}
cols_to_delete <- 11:(ncol(picks_wide) - 1)
do_not_show_list <-  as.list(cols_to_delete)

for (i in seq_along(do_not_show_list)) {
  names(do_not_show_list[[i]]) <- "targets"
}


for(i in (cols_to_delete - 10) ) {
  #  do_not_show_list[[i]]$targets <- as.list(cols_to_delete)[[i]]
  suppressWarnings(do_not_show_list[[i]]$visible <- FALSE)
}

point_levels <- rep(seq(15, 6, by = -1), 2) %>% 
  sort(decreasing = TRUE)
paste_levels <- point_levels %>% 
  paste(c(knitr::asis_output("\U2193"), knitr::asis_output("\U2191")))

picks_wide %>% 
  # Doesn't actually sort in DT
  mutate_at(
    .vars = c("Adonis", "Chester", "Jake", "Jenelle",
              "Mary", "Mike", "Phil", "Ryan"), 
    .funs = ~ factor(.x, levels = paste_levels)) %>% 
  #  purrr::map_df(rev) %>% 
  arrange(Team) %>% 
  DT::datatable(rownames = FALSE, options = list(
    pageLength = 30, dom = 'ft',
    columnDefs = do_not_show_list
    #list(list(targets = as.list(cols_to_delete), visible = FALSE))
  )) %>% 
  formatStyle(
    "Chester", "chester_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Jake", "jake_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Jenelle", "jenelle_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Adonis", "adonis_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Mike", "mike_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Mary", "mary_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Ryan", "ryan_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Phil", "phil_correct",
    target = 'cell',
    backgroundColor = styleEqual(c(TRUE, FALSE, NA), 
                                 c("lightgreen", "hotpink", "white"))
  ) %>% 
  formatStyle(
    "Team", target = 'cell', backgroundColor = "white"
  ) %>% 
  formatStyle(
    "Win Target", target = 'cell', backgroundColor = "white"
  )  %>% 
  formatStyle(
    "Current Wins", target = 'cell', backgroundColor = "white"
  )  
```

***

#### Point explanations by team for `r Sys.Date() - 1` {.tabset}

```{r message=FALSE}
library(DT)
player_table <- function(player_num) {
  df <- player_projections_by_team %>% 
    filter(Player == current_rankings[player_num],
           Date == Sys.Date() - 1) %>% 
    arrange(desc(`Current Projected Points`)) %>% 
    select(-Player, -Date)
  
  # if (sum(is.na(df$`Outcome Determined`)) == 30) {
  #   df %>% 
  #     select(-`Outcome Determined`) %>% 
  #     datatable(options = list(pageLength = 30, dom = 'ft'))
  # } else {
  df %>%
    datatable(rownames = FALSE,
              extensions = "FixedColumns",
              options = list(pageLength = 30, dom = 'ft',
                             fixedColumns = list(leftColumns = 2))
    ) # %>%
  # formatStyle(
  #   "Outcome Determined",
  #   target = 'row',
  #   backgroundColor = styleEqual(c("OVER", "UNDER", NA_character_), 
  #                                c("lightgreen", "deeppink", "white"))
  # )
  # }
  
  # brks <- quantile(
  #   df$`Wins To Go Over Vegas Insider`, 
  #   probs = seq(.05, .95, 0.1), 
  #   na.rm = TRUE
  # )
  # clrs <- rev(brewer.pal(n = 11, name = "PiYG"))
  # 
  # df %>% 
  #   datatable(options = list(pageLength = 30, dom = 'ft')) %>% 
  #   formatStyle(
  #     'Wins To Go Over Vegas Insider',
  #     target = 'row',
  #     backgroundColor = styleInterval(brks, clrs)
  #   )
}
```


##### `r current_rankings[1]`

```{r}
player_table(1)
```

##### `r current_rankings[2]`

```{r}
player_table(2)
```

##### `r current_rankings[3]`

```{r}
player_table(3)
```

##### `r current_rankings[4]`

```{r}
player_table(4)
```

##### `r current_rankings[5]`

```{r}
player_table(5)
```

##### `r current_rankings[6]`

```{r}
player_table(6)
```

##### `r current_rankings[7]`

```{r}
player_table(7)
```

##### `r current_rankings[8]`

```{r}
player_table(8)
```


***

### All together

```{r fig.height=6, fig.width=10}
projected_score_highlight <- projected_score %>% 
  mutate(Player = factor(x = Player, levels = current_rankings)) %>%
  highlight_key(~Player)
projected_score_plot <- projected_score_highlight %>% 
  ggplot(aes(
    x = Date, y = `Projected Total Points`, color = Player)
  ) +
  geom_point() +
  geom_line() +
  xlab("") + 
  ylab("") + 
  scale_color_brewer(palette = "RdYlBu") +
  theme_bw()

gg <- ggplotly(projected_score_plot, 
               tooltip = c("Date", "Projected Total Points", "Player")) %>% 
  layout(
    margin = list(t = 80),
    title = list(
      text = "Projected points over time",
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c(rep("&nbsp;", 15), "Projected Total Points",  
          "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 9), "Date", "<br>"), collapse = "")
    ))
highlight(gg, color = c("lightgreen"), 
          off = "plotly_doubleclick",
          #          on = "plotly_hover", off = "plotly_deselect", 
          dynamic = FALSE)
```

### Small multiples

```{r fig.height=6, fig.width=10}
projected_score_facetted <- projected_score %>% 
  mutate(Player = factor(x = Player, levels = current_rankings)) %>% 
  ggplot(aes(
    x = Date, y = `Projected Total Points`)
  ) +
  geom_point() +
  geom_line() +
  xlab("") + 
  ylab("") + 
  theme_bw() +
  facet_wrap(~ Player, nrow = 2)

ggplotly(projected_score_facetted) %>% 
  layout(
    margin = list(t = 80),
    title = list(
      text = "Projected points (sorted from current best to worst) over time",
      font = list(size = 20)
    ),
    yaxis = list(
      title = paste(
        c("Projected Total Points", rep("&nbsp;", 30), "<br>"), collapse = "")
    ),
    xaxis = list(
      title = paste(
        c(rep("&nbsp;", 105), "Date", "<br>"), collapse = "")
    ))
```



## Changes in projected points since yesterday

```{r results="asis"}
if (nrow(rows_today) == 0) {
  cat(paste("There was no change in scores for all players based on the",
            "outcomes of yesterday's games."))
}
```

```{r}
if (nrow(rows_today)) {
  changes_in_player_projections %>% 
    arrange(desc(Date), Team, Player) %>% 
    DT::datatable(options = list(pageLength = nrow(rows_today)))
}
```

```{r}
yesterday_scores <- scores %>% 
  filter(game_date == Sys.Date() - 1) %>% 
  select(`Game Date` = game_date,
         `Away Team` = away_team,
         `Away Score` = away_score,
         `Home Team` = home_team,
         `Home Score` = home_score)
```

***

```{r results="asis"}
if (nrow(yesterday_scores)) {
  cat("Yesterday's scores")
} else {
  cat("No NBA games were played yesterday.")
}
```

```{r}
if (nrow(yesterday_scores)) {
  yesterday_scores %>% knitr::kable()
  #  DT::datatable(options = list(pageLength = nrow(yesterday_scores)))
}
```


```{r}
schedule <- nbastatR::current_schedule() %>% 
  filter(dateGame >= "2020-12-22") %>% 
  select(game_date = dateGame,
         game_id = idGame,
         slug_away_team = slugTeamAway,
         away_team = nameTeamAway,
         slug_home_team = slugTeamHome,
         home_team = nameTeamHome,
         start_time = hasBuzzerBeater
  )
todays_games <- schedule %>% 
  filter(game_date == Sys.Date()) %>% 
  select(`Game Date` = game_date,
         `Start Time` = start_time,
         `Away Team` = away_team,
         `Home Team` = home_team)
```

***

```{r results="asis"}
if (nrow(todays_games)) {
  cat("Today's games")
} else {
  cat("No NBA games are scheduled for today.")
}
```

```{r}
if (nrow(todays_games)) {
  todays_games %>% knitr::kable()
  #  DT::datatable(options = list(pageLength = nrow(yesterday_scores)))
}
```

## Current standings {.tabset}

### By conference {.tabset}

Note that tie breakers to determine playoff seeds are not currently set.

#### Western

```{r}
standings %>% 
  filter(conference == "West") %>% 
  select(-conference) %>% 
  relocate(division, .after = team_name) %>% 
  arrange(desc(differential)) %>% 
  select(-differential) %>% 
  rename(
    Team = team_name, 
    Wins = wins, 
    Losses = losses, 
    Division = division
  ) %>% 
  knitr::kable(align = c('l', 'l', 'r', 'r', 'r', 'r'))
```

#### Eastern

```{r}
standings %>% 
  filter(conference == "East") %>% 
  select(-conference) %>% 
  relocate(division, .after = team_name) %>% 
  arrange(desc(differential)) %>% 
  select(-differential) %>% 
  rename(
    Team = team_name, 
    Wins = wins, 
    Losses = losses, 
    Division = division
  ) %>% 
  knitr::kable(align = c('l', 'l', 'r', 'r', 'r', 'r'))
```


### By division {.tabset}

#### West {.tabset}

##### Pacific

```{r}
standings %>% 
  filter(division == "Pacific") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable()
```

##### Northwest

```{r}
standings %>% 
  filter(division == "Northwest") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable()
```

##### Southwest

```{r}
standings %>% 
  filter(division == "Southwest") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable()
```

#### East {.tabset}

##### Central

```{r}
standings %>% 
  filter(division == "Central") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable()
```

##### Atlantic

```{r}
standings %>% 
  filter(division == "Atlantic") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable()
```

##### Southeast

```{r}
standings %>% 
  filter(division == "Southeast") %>% 
  select(-conference, -division) %>% 
  arrange(desc(`Winning Pct`)) %>% 
  select(Team = team_name, Wins = wins, Losses = losses, `Winning Pct`) %>% 
  knitr::kable()
```
